---
title: "Narrow_Down"
author: "Carlos Haring"
date: "2025-06-26"
output: html_document
---

```{r setup, include=FALSE}
clinData <- read.csv("https://raw.githubusercontent.com/BDSI-CDS-2025/tcia/refs/heads/main/data/clinicalData_clean.csv")
```

Load packages that are necessary for running all code below.
```{r}
#' example code for installing packages for first time:
#' install.packages("tidyverse")

library(tidyverse)
library(ggplot2)
library(survival)
library(ggfortify)
library(devtools)
#' install_github("jokergoo/ComplexHeatmap")
library(ComplexHeatmap)
library(umap)
library(gtsummary)
library(summarytools)
library(flextable)
# 
install.packages("Visdot")  
library(Visdot)
install.packages("installr")
library(installr)
updateR()
```


## Reading Clinical Data

```{r}
clinData <- clinData[,-1] # remove first column of row numbers since we have
# patient ID
```

## Data cleaning 

### changing race

```{r}
clinData <- clinData %>%
  mutate(Race.and.Ethnicity = case_when(Race.and.Ethnicity == 1 ~ "White", Race.and.Ethnicity %in% c(2 ,3, 4, 5, 6, 7) ~ "Other"))

```

### recoding molecular subtype

```{r}
clinData <- clinData %>% 
mutate(Mol.Subtype. = case_when(Mol.Subtype == 0 ~ "luminal-like", Mol.Subtype == 1 ~ "ER/PR pos, HER2 pos", Mol.Subtype == 2 ~ "HER2 pos", Mol.Subtype == 3 ~ "triple neg"))
```

### recoding Metastatic at presentation

```{r}
clinData <- clinData %>% 
mutate(Metastatic.at.Presentation..Outside.of.Lymph.Nodes. = case_when(Metastatic.at.Presentation..Outside.of.Lymph.Nodes. == 0 ~ "No", Metastatic.at.Presentation..Outside.of.Lymph.Nodes. == 1 ~ "Yes"))
```

### recoding race

```{r}
clinData <- clinData %>%
  mutate(Recurrence.event.s.clean = as.factor(case_when(
      Recurrence.event.s. == 1 ~ "Recurrence",
      Recurrence.event.s. == 0 ~ "No Recurrence"))) %>%
  filter(!is.na(Recurrence.event.s.clean))
  
```

### recoding  reaccuance 

```{r}
clinData <- clinData %>%
  mutate(
    Days.to.local.recurrence = as.numeric(na_if(Days.to.local.recurrence..from.the.date.of.diagnosis., "NP")),
    Days.to.distant.recurrence = as.numeric(na_if(Days.to.distant.recurrence.from.the.date.of.diagnosis., "NP")),
    Days.to.last.recurrence.free = as.numeric(na_if(Days.to.last.local.recurrence.free.assessment..from.the.date.of.diagnosis., "NP"))
  ) %>%
  mutate(
    time.to.recurrence = case_when(
      !is.na(Days.to.local.recurrence) & !is.na(Days.to.distant.recurrence) ~
        pmin(Days.to.local.recurrence, Days.to.distant.recurrence, na.rm = TRUE),
      
      !is.na(Days.to.local.recurrence) ~ Days.to.local.recurrence,
      
      !is.na(Days.to.distant.recurrence) ~ Days.to.distant.recurrence,
      
      TRUE ~ Days.to.last.recurrence.free  
    )
  )
```

### recoding size

```{r}
clinData <- clinData %>%
  mutate(
    TumorSize_mammo = as.numeric(na_if(na_if(TumorSize_mammo, "NC"), "NA")),
    TumorSize_us = as.numeric(na_if(na_if(TumorSize_us, "NC"), "NA"))) %>%
  rowwise() %>%
  mutate(
    size = case_when(
      !is.na(TumorSize_mammo) & !is.na(TumorSize_us) ~
        mean(c_across(c(TumorSize_mammo, TumorSize_us)), na.rm = TRUE),
      
      !is.na(TumorSize_mammo) ~ TumorSize_mammo,
      
      !is.na(TumorSize_us) ~ TumorSize_us,
      
      TRUE ~ NA_real_  
    )
  ) %>%
  ungroup()

```

### Anti her2 column

```{r}
clinData <- clinData %>%
  mutate(Anti.Her2.Neu.Therapy = case_when(Neoadjuvant.Anti.Her2.Neu.Therapy == 1 &  Adjuvant.Anti.Her2.Neu.Therapy == 1 ~ "Her2 Therapy",
    Neoadjuvant.Anti.Her2.Neu.Therapy == 1 ~ "Her2 Therapy", 
    Adjuvant.Anti.Her2.Neu.Therapy == 1 ~ "Her2 Therapy", 
    Neoadjuvant.Anti.Her2.Neu.Therapy == 0 ~ "No Her2 Therapy", 
    Adjuvant.Anti.Her2.Neu.Therapy == 0 ~ "No Her2 Therapy"))
```

### Endocrine therapy column

```{r}
clinData <- clinData %>%
   mutate(Endocrine.Therapy.Medication = case_when(Neoadjuvant.Endocrine.Therapy.Medications == 1 ~ "Endocrine Therapy", 
                                                   Adjuvant.Endocrine.Therapy.Medications == 1 ~ "Endocrine Therapy", 
                                                   Neoadjuvant.Endocrine.Therapy.Medications == 0 ~ " No Endocrine Therapy", 
                                                   Adjuvant.Endocrine.Therapy.Medications == 0 ~ "No Endocrine Therapy"))
```

### Radiation therapy column

```{r}
clinData <- clinData %>%
  mutate(Chemotherapy = case_when(Adjuvant.Chemotherapy == 1 ~ "Chemotherapy", 
                                  Neoadjuvant.Chemotherapy == 1 ~ "Chemotherapy", 
                                  Adjuvant.Chemotherapy == 0 ~ "No Chemotherapy", 
                                  Adjuvant.Chemotherapy == 0~ "No Chemotherapy"))
```

### chemotherapy column

```{r}
clinData <- clinData %>%
  mutate(Radiation.Therapy = case_when(Adjuvant.Radiation.Therapy == 1 ~ "Radiation Therapy", 
                                Neoadjuvant.Endocrine.Therapy.Medications == 1 ~"Radiation Therapy", 
                                Adjuvant.Radiation.Therapy == 0~ "No Radiation Therapy", 
                                Neoadjuvant.Endocrine.Therapy.Medications == 0 ~"No Radiation Therapy"))
```

### ovaries in situ 

```{r}
clinData <- clinData %>%
  mutate(Number.of.Ovaries.In.Situ.. = case_when(Number.of.Ovaries.In.Situ.. %in% c(1, 2) ~ "Intact Ovaries ≥ 1", 
                                Number.of.Ovaries.In.Situ.. == 0 ~"No Ovaries Intact", 
 
                               Number.of.Ovaries.In.Situ.. == "NP" ~ NA_character_))
```

### pathologic response T

```{r}
clinData <- clinData %>%
  mutate(Pathologic.response.T = case_when(Pathologic.response.to.Neoadjuvant.therapy..Pathologic.stage..T..following.neoadjuvant.therapy %in% c("1", "0") ~ "Small/ no Tumor remains", 
                                Pathologic.response.to.Neoadjuvant.therapy..Pathologic.stage..T..following.neoadjuvant.therapy %in% c("2", "3", "4", "5") ~ "DCIS or ≥ medium tumor", 
 
                               Pathologic.response.to.Neoadjuvant.therapy..Pathologic.stage..T..following.neoadjuvant.therapy %in% c("NA", "-1") ~ NA_character_))
```

### pathologic response N

```{r}
clinData <- clinData %>%
  mutate(Pathologic.response.N = case_when(Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..N..following.neoadjuvant.therapy ==  "0" ~ "Cancer in not lymph", 
                                Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..N..following.neoadjuvant.therapy %in% c("1", "2", "3") ~ "Cancer in not lymph", 
                               Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..N..following.neoadjuvant.therapy %in% c("NA", "-1") ~ NA_character_))
```

### Tumor size, staging nodes, staging metastasis

```{r}

clinData <- clinData %>%
  mutate(
    Staging.Tumor.Size....T. = factor(Staging.Tumor.Size....T.,
                     levels = c(1, 2, 3, 4),
                     ordered = TRUE),

    Staging.Nodes...Nx.replaced.by..1..N. = factor(na_if(Staging.Nodes...Nx.replaced.by..1..N., -1),
                     levels = c(0, 1, 2, 3),
                     ordered = TRUE),

    Staging.Metastasis...Mx..replaced.by..1..M. = factor(na_if(Staging.Metastasis...Mx..replaced.by..1..M., -1),
                     levels = c(0, 1),
                     ordered = TRUE)
  )

```

### mole subtype

```{r}
clinData <- clinData %>% 
  mutate(Mol.Subtype = as.factor(Mol.Subtype))
```

### Tumor Size, Stady Nodes, and Metastasis

```{r}
clinData <- clinData %>%
  mutate(
    Staging.Tumor.Size....T. = factor(Staging.Tumor.Size....T., 
                                      levels = c(1, 2, 3, 4), 
                                      ordered = TRUE),

    Staging.Nodes...Nx.replaced.by..1..N. = factor(
      na_if(as.numeric(as.character(Staging.Nodes...Nx.replaced.by..1..N.)), -1),
      levels = c(0, 1, 2, 3),
      ordered = TRUE
    ),

    Staging.Metastasis...Mx..replaced.by..1..M. = factor(
      na_if(as.numeric(as.character(Staging.Metastasis...Mx..replaced.by..1..M.)), -1),
      levels = c(0, 1),
      ordered = TRUE
    )
  )

```

### Nottinham_total_score

```{r}
clinData <- clinData %>%
  mutate(
    Nottingham_total_score = as.numeric(TumorGradeT) + as.numeric(TumorGradeN) + as.numeric(TumorGradeM)
  )
```

### Histologic Type 

```{r}
clinData$Histologic.type.new <- fct_collapse(
  factor(clinData$Histologic.type),
  Ductal = c("1"),
  Other = c("0","2", "3", "4", "5", "6", "7", "8", "9", "10")
)
```

### Bilateral Information

```{r}
clinData <- clinData %>% 
  mutate(Bilateral.Information = as.factor(na_if(Bilateral.Information, "NC")))
```

### Definative surgery type 

```{r}
clinData <- clinData %>%
  mutate(
    Surgery = as.factor(Surgery),
    Definitive.Surgery.Type = as.factor(na_if(Definitive.Surgery.Type, "NP")),
    Days.to.Surgery..from.the.date.of.diagnosis. = as.numeric(Days.to.Surgery..from.the.date.of.diagnosis.),
    
    Days.to.Surgery = ifelse(
      Days.to.Surgery..from.the.date.of.diagnosis. < 0, 
      NA,  
      Days.to.Surgery..from.the.date.of.diagnosis.
    )
  )

```

### Time To Recurrence 

```{r}
clinData <- clinData %>%
  mutate(
    time.to.recurrence.new = ifelse(
      !is.na(Days.to.Surgery) & !is.na(time.to.recurrence),
      time.to.recurrence - Days.to.Surgery,
      NA_real_
    )
  )
```

## Filtering by significant variables 

```{r}
clinData_sig <- clinData %>%
   select(c("Definitive.Surgery.Type", "Nottingham.grade", "Staging.Tumor.Size....T.", "Staging.Nodes...Nx.replaced.by..1..N.", "Mol.Subtype", "Race.and.Ethnicity", "Metastatic.at.Presentation..Outside.of.Lymph.Nodes.", "Neoadjuvant.Chemotherapy", "Chemotherapy", "Adjuvant.Endocrine.Therapy.Medications", "Pathologic.response.T", "Received.Neoadjuvant.Therapy.or.Not", "Endocrine.Therapy.Medication"))
```

## 

```{r}

```

