---
title: "Carlos_prelim_anaylsis"
author: "Carlos Haring"
date: "2025-06-18"
output: html_document
---

```{r setup, include=FALSE}
cdc <- read.csv("https://raw.githubusercontent.com/BDSI-CDS-2025/tcia/refs/heads/main/data/clinicalData_clean.csv")
img <- read.csv("https://raw.githubusercontent.com/BDSI-CDS-2025/tcia/refs/heads/Carlos-Branch/data/imagingFeatures.csv")
```

Load packages that are necessary for running all code below.
```{r}
#' example code for installing packages for first time:
#' install.packages("tidyverse")

library(tidyverse)
library(ggplot2)
library(survival)
library(ggfortify)
library(devtools)
#' install_github("jokergoo/ComplexHeatmap")
library(ComplexHeatmap)
library(umap)
library(gtsummary)
library(summarytools)
library(flextable)
```


## Reading Clinical Data

```{r}
clinData <- cdc
clinData <- clinData[,-1] # remove first column of row numbers since we have
# patient ID
```

The '<-' is the assignment operator in R. clinData <- read.csv(...) assigns 
the output of read.csv() into a variable/object called clinData. Keyboard 
shortcut on a Mac is option, -. You can also use '='. '<-' is used to
differentiate setting function arguments versus assigning data to a variable.

## Reading Image Features
```{r}
imFeatures <- img
imFeatures <- imFeatures[,-1]
```

In general, the featureCitations.docx file describes the features and provides 
citations. The features are grouped into the categories listed below.

Breast and fibroglandular tissue (FGT) volume features: volume and density of 
breast area and fibroglandular area.

Tumor size and morphology: regularity/roundness vs irregularity of tumor
shape and size.

FGT enhancement: measure how much FGT is enhanced when we add contrast.
This is referred to as BPE, background parenchymal enhancement, which may 
confound tumor enhancement/identification.

Tumor enhancement: measure how much tumor is enhanced when we add contrast

Combining tumor and FGT enhancement: measure how tumor and FGT are 
enhanced when adding contrast

FGT enhancement texture: describe the enhancement due to contrast. Looking at
local patterns, does enhancement look gritty or smooth?

Tumor enhancement texture: same as FGT but for tumor area.

Tumor Enhancement Spatial Heterogeneity: measure similarity between tumor
subregions of the tumor.

FGT enhancement variation: global variation of contrast enhancement of FGT

Tumor enhancement variation: global variation of contrast enhancement of tumor

## Task 1: What is in the data?

How many patients are in the dataset?

```{r}
nrow(clinData)
#922 patients
```


How many missing values are there among image features?

```{r}
sum(is.na(imFeatures))

missingness <- apply(
  imFeatures, 2, function(x){
    sum(is.na(x))
  }
)
sum(missingness == 0) # there are 490 variables with no missingness
sum(missingness != 0) # the remaining 40 variables have some missing values
missingness[missingness != 0]
```

How many missing values are there among clinical features?

```{r}
sum(is.na(clinData))

sum(missingness == 0) #490 has no missing values
sum(missigness =! 0) #only one missing value
```


What types of variables are in the clinical dataset? Are they quantitative?

```{r}
str(clinData)
clinData %>%
  select(where(is.character)) #36 character column

clinData %>%
  select(where(is.integer)) #59 quantitative

clinData %>%
  select(where(is.factor)) #no factors

clinData %>%
  select(where(is.numeric)) #61 are numeric

ncol(clinData)
```

## Data cleaning 

### changing race

```{r}
clinData <- clinData %>%
  mutate(Race.and.Ethnicity = case_when(Race.and.Ethnicity == 1 ~ "White", Race.and.Ethnicity %in% c(2 ,3, 4, 5, 6, 7) ~ "Other"))

```

### recoding molecular subtype

```{r}
clinData <- clinData %>% 
mutate(Mol.Subtype. = case_when(Mol.Subtype == 0 ~ "luminal-like", Mol.Subtype == 1 ~ "ER/PR pos, HER2 pos", Mol.Subtype == 2 ~ "HER2 pos", Mol.Subtype == 3 ~ "triple neg"))
```

### recoding Metastatic at presentation

```{r}
clinData <- clinData %>% 
mutate(Metastatic.at.Presentation..Outside.of.Lymph.Nodes. = case_when(Metastatic.at.Presentation..Outside.of.Lymph.Nodes. == 0 ~ "No", Metastatic.at.Presentation..Outside.of.Lymph.Nodes. == 1 ~ "Yes"))
```

### recoding race

```{r}
clinData <- clinData %>%
  mutate(Recurrence.event.s.clean = as.factor(case_when(
      Recurrence.event.s. == 1 ~ "Recurrence",
      Recurrence.event.s. == 0 ~ "No Recurrence"))) %>%
  filter(!is.na(Recurrence.event.s.clean))
  
```

### recoding  reaccuance 

```{r}
clinData <- clinData %>%
  mutate(
    Days.to.local.recurrence = as.numeric(na_if(Days.to.local.recurrence..from.the.date.of.diagnosis., "NP")),
    Days.to.distant.recurrence = as.numeric(na_if(Days.to.distant.recurrence.from.the.date.of.diagnosis., "NP")),
    Days.to.last.recurrence.free = as.numeric(na_if(Days.to.last.local.recurrence.free.assessment..from.the.date.of.diagnosis., "NP"))
  ) %>%
  mutate(
    time.to.recurrence = case_when(
      !is.na(Days.to.local.recurrence) & !is.na(Days.to.distant.recurrence) ~
        pmin(Days.to.local.recurrence, Days.to.distant.recurrence, na.rm = TRUE),
      
      !is.na(Days.to.local.recurrence) ~ Days.to.local.recurrence,
      
      !is.na(Days.to.distant.recurrence) ~ Days.to.distant.recurrence,
      
      TRUE ~ Days.to.last.recurrence.free  
    )
  )
```

### recoding size

```{r}
clinData <- clinData %>%
  mutate(
    TumorSize_mammo = as.numeric(na_if(na_if(TumorSize_mammo, "NC"), "NA")),
    TumorSize_us = as.numeric(na_if(na_if(TumorSize_us, "NC"), "NA"))) %>%
  rowwise() %>%
  mutate(
    size = case_when(
      !is.na(TumorSize_mammo) & !is.na(TumorSize_us) ~
        mean(c_across(c(TumorSize_mammo, TumorSize_us)), na.rm = TRUE),
      
      !is.na(TumorSize_mammo) ~ TumorSize_mammo,
      
      !is.na(TumorSize_us) ~ TumorSize_us,
      
      TRUE ~ NA_real_  
    )
  ) %>%
  ungroup()

```

### Anti her2 column

```{r}
clinData <- clinData %>%
  mutate(Anti.Her2.Neu.Therapy = case_when(Neoadjuvant.Anti.Her2.Neu.Therapy == 1 &  Adjuvant.Anti.Her2.Neu.Therapy == 1 ~ "Her2 Therapy",
    Neoadjuvant.Anti.Her2.Neu.Therapy == 1 ~ "Her2 Therapy", 
    Adjuvant.Anti.Her2.Neu.Therapy == 1 ~ "Her2 Therapy", 
    Neoadjuvant.Anti.Her2.Neu.Therapy == 0 ~ "No Her2 Therapy", 
    Adjuvant.Anti.Her2.Neu.Therapy == 0 ~ "No Her2 Therapy"))
```

### Endocrine therapy column

```{r}
clinData <- clinData %>%
   mutate(Endocrine.Therapy.Medication = case_when(Neoadjuvant.Endocrine.Therapy.Medications == 1 ~ "Endocrine Therapy", 
                                                   Adjuvant.Endocrine.Therapy.Medications == 1 ~ "Endocrine Therapy", 
                                                   Neoadjuvant.Endocrine.Therapy.Medications == 0 ~ " No Endocrine Therapy", 
                                                   Adjuvant.Endocrine.Therapy.Medications == 0 ~ "No Endocrine Therapy"))
```

### Radiation therapy column

```{r}
clinData <- clinData %>%
  mutate(Chemotherapy = case_when(Adjuvant.Chemotherapy == 1 ~ "Chemotherapy", 
                                  Neoadjuvant.Chemotherapy == 1 ~ "Chemotherapy", 
                                  Adjuvant.Chemotherapy == 0 ~ "No Chemotherapy", 
                                  Adjuvant.Chemotherapy == 0~ "No Chemotherapy"))
```

### chemotherapy column

```{r}
clinData <- clinData %>%
  mutate(Radiation.Therapy = case_when(Adjuvant.Radiation.Therapy == 1 ~ "Radiation Therapy", 
                                Neoadjuvant.Endocrine.Therapy.Medications == 1 ~"Radiation Therapy", 
                                Adjuvant.Radiation.Therapy == 0~ "No Radiation Therapy", 
                                Neoadjuvant.Endocrine.Therapy.Medications == 0 ~"No Radiation Therapy"))
```

### ovaries in situ 

```{r}
clinData <- clinData %>%
  mutate(Number.of.Ovaries.In.Situ.. = case_when(Number.of.Ovaries.In.Situ.. %in% c(1, 2) ~ "Intact Ovaries ≥ 1", 
                                Number.of.Ovaries.In.Situ.. == 0 ~"No Ovaries Intact", 
 
                               Number.of.Ovaries.In.Situ.. == "NP" ~ NA_character_))
```

### pathologic response T

```{r}
clinData <- clinData %>%
  mutate(Pathologic.response.T = case_when(Pathologic.response.to.Neoadjuvant.therapy..Pathologic.stage..T..following.neoadjuvant.therapy %in% c("1", "0") ~ "Small/ no Tumor remains", 
                                Pathologic.response.to.Neoadjuvant.therapy..Pathologic.stage..T..following.neoadjuvant.therapy %in% c("2", "3", "4", "5") ~ "DCIS or ≥ medium tumor", 
 
                               Pathologic.response.to.Neoadjuvant.therapy..Pathologic.stage..T..following.neoadjuvant.therapy %in% c("NA", "-1") ~ NA_character_))
```

### pathologic response N

```{r}
clinData <- clinData %>%
  mutate(Pathologic.response.N = case_when(Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..N..following.neoadjuvant.therapy ==  "0" ~ "Cancer in not lymph", 
                                Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..N..following.neoadjuvant.therapy %in% c("1", "2", "3") ~ "Cancer not lymph", 
                               Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..N..following.neoadjuvant.therapy %in% c("NA", "-1") ~ NA_character_))
```


# Data exploration 

## exploring non binonial variables 

### N response 

```{r}
table(clinData$Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..N..following.neoadjuvant.therapy, clinData$Recurrence.event.s.) # usable but needs combination 
```
### pathologic response M

```{r}
table(clinData$Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..M..following.neoadjuvant.therapy, clinData$Recurrence.event.s.) # not usable 
```
### near complete response looser

```{r}
table(clinData$Overall.Near.complete.Response...Looser.Definition, clinData$Recurrence.event.s.) # not usable 

```
### near complete response stricter

```{r}
table(clinData$Overall.Near.complete.Response...Stricter.Definition, clinData$Recurrence.event.s.) # not usable 
```
### near complete response stricter

```{r}
table(clinData$Near.complete.Response..Graded.Measure., clinData$Recurrence.event.s.) # not usable 
```

### tumor response investigation

```{r}
table(clinData$Clinical.Response..Evaluated.Through.Imaging, clinData$Recurrence.event.s.) # clinical response not usable 
```

### Patholigc response to Neoadjuvent Therapy investigation

```{r}
table(clinData$Pathologic.Response.to.Neoadjuvant.Therapy, clinData$Recurrence.event.s.) # Pathologic response not usable 
```

### Breast density 

```{r}
clinData_clean <- clinData %>% filter(!is.na(Age.at.mammo..days.), !is.na(time.to.recurrence))
ggplot(clinData_clean, aes(x = Age.at.mammo..days., y = time.to.recurrence, color = Recurrence.event.s.)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x)
```

### breast density

```{r}
table(clinData$Breast.Density, clinData$Recurrence.event.s.) # breast not usable 

```

### breast shape

```{r}
table(clinData$Margin_mammo, clinData$Recurrence.event.s.) # not usable
```
### mass density

```{r}
table(clinData$Mass.Density, clinData$Recurrence.event.s.) # not usable
```

### calcifications

```{r}
table(clinData$Calcifications, clinData$Recurrence.event.s.) # not usable
```

### tumor size us
```{r}
table(clinData$TumorSize_us, clinData$Recurrence.event.s.) # not usable
```

### margin us 

```{r}
table(clinData$Recurrence.event.s., clinData$Margin_us) # not usable
```

### margin us 

```{r}
table(clinData$Echogenicity, clinData$Recurrence.event.s.) # not usable)
```

### margin ovary number in situ 

```{r}
table(clinData$Number.of.Ovaries.In.Situ.., clinData$Recurrence.event.s.) #was non binomial now is binomial usable 
```
### margin ovary number in situ 

```{r}
table(clinData$Pathologic.response.to.Neoadjuvant.therapy..Pathologic.stage..T..following.neoadjuvant.therapy, clinData$Recurrence.event.s.) #was non binomial now is binomial usable 
```

## Exploring the general outcome

### Day to local reacurrance

```{r}
hist(clinData$Days.to.local.recurrence)
summarytools::dfSummary(clinData$Days.to.local.recurrence)
```

### Day to distant reacurrance

```{r}
hist(clinData$Days.to.distant.recurrence)
summarytools::dfSummary(clinData$Days.to.distant.recurrence)

```

### Day to first reaccurance

```{r}
hist(clinData$time.to.recurrence)
summarytools::dfSummary(clinData$time.to.recurrence)
```
### table for 

```{r}
# smaller_df <- clinData %>% 
#   tbl_summary(by = Recurrence.event.s., #by segregates our graph by Recurrence type
#                       label = list(
#                                   # labels each thing
#                                    Race.and.Ethnicity.clean = "Race/Ethnicity",
#                                    Mol.Subtype.clean = "Cancer Molecular Subtype",
#                                    Surgery.clean = "Occurance of Surgery",
#                                    Metastatic.at.Presentation..Outside.of.Lymph.Nodes.clean = "Metastatic Outside of Lymph Node")) %>%
#   add_p() %>%  as_flex_table() %>%
#   width(j = 1, width = 2, unit = "in") %>%
#   width(j = 2:4, width = 1.3, unit = "in")
```

# determining which variables to check with kaplian myer

## filter the variables i am responsible for

```{r}
car_clinData <- clinData[, 57:109]
```

## function for getting colnamees of functions with greater than 9 events

```{r}
valid_columns <- c()
invalid_columns <- c()
check_contingency_counts <- function(data, outcome_col = "Recurrence.event.s.") {
    valid_columns <- c()
  invalid_columns <- c()
  for (colname in colnames(data)) {
    if (colname == outcome_col) next 
        tab <- table(data[[colname]], data[[outcome_col]])
        if (all(tab >= 9)) {
      valid_columns <- c(valid_columns, colname) 
        } else {
      invalid_columns <- c(invalid_columns, colname) 
        }
  }
   return(list(valid = valid_columns, invalid = invalid_columns))
}
```

## using said function 

```{r}
list <- check_contingency_counts(car_clinData)
list$valid
```

# Kaplin Myers EDA

## filter valid column

```{r}
car_clinData_valid <- car_clinData[ , c(list$valid, "time.to.recurrence", "Recurrence.event.s.")]
car_clinData_valid

kmFit_subtype <- survfit(Surv(time = clinData[["time.to.recurrence"]], event = clinData[["Recurrence.event.s."]]) ~ clinData[["Pathologic.response.N"]], data = clinData)
the_plot <- autoplot(kmFit_subtype)
the_plot


survdiff(Surv_obj~kmFit_subtype,data = clinData)

```

## function for creating Kaplian-Mayer

```{r}

Kap_creator_significant <- function(the_data, the_event = "Recurrence.event.s.", the_time = "time.to.recurrence") {
  plot_list <- list()
  
  for (colname in colnames(the_data)) {
     if (colname == "Recurrence.event.s.") next 
     if (colname == "time.to.recurrence") next 
ttle <- paste0("Time to Reaccurance by ",colname)
kmFit_subtype <- survfit(Surv(time = the_data[[the_time]], event = the_data[[the_event]]) ~ the_data[[colname]], data = the_data)
the_plot <- autoplot(kmFit_subtype) +
  labs(col = ttle, fill = ttle, title = ttle) +  theme_light() + theme(legend.position = "bottom") + xlab("Time days") + ylab("Recurrence Free Probability")
    plot_list[[colname]] <- the_plot
    plotname <- paste0(colname,".png")
    ggsave(plotname, the_plot, path = "/Users/carlosh/Documents/Michigan_stuff/project/Survival_rate/Recur_Surv/Carlos_KM_plot/Sig_plot", width = 8, height = 5)
  }
  return(plot_list)
  
}

Kap_creator <- function(the_data, the_event = "Recurrence.event.s.", the_time = "time.to.recurrence") {
  plot_list <- list()
  
  for (colname in colnames(the_data)) {
     if (colname == "Recurrence.event.s.") next 
     if (colname == "time.to.recurrence") next 
ttle <- paste0("Time to Reaccurance by ",colname)
kmFit_subtype <- survfit(Surv(time = the_data[[the_time]], event = the_data[[the_event]]) ~ the_data[[colname]], data = the_data)
the_plot <- autoplot(kmFit_subtype) +
  labs(col = ttle, fill = ttle, title = ttle) +  theme_light() + theme(legend.position = "bottom") + xlab("Time days") + ylab("Recurrence Free Probability")
    plot_list[[colname]] <- the_plot
    plotname <- paste0(colname,".png")
    ggsave(plotname, the_plot, path = "/Users/carlosh/Documents/Michigan_stuff/project/Survival_rate/Recur_Surv/Carlos_KM_plot")
  }
  return(plot_list)
  
}
```

## Checking if they work

```{r}

Kap_creator(car_clinData_valid)

```
car_clinData_valid <- car_clinData[ , c(list$valid, "time.to.recurrence", "Recurrence.event.s.")]
car_clinData_valid


## singificant km creator

```{r}
car_clinData_valid <- car_clinData[ , c(sig_list, "time.to.recurrence", "Recurrence.event.s.")]
Kap_creator(car_clinData_valid)
```


## function for running the Log rank test 

```{r}

log_ranker <- function(the_data, the_event = "Recurrence.event.s.", the_time = "time.to.recurrence") {
  sig_list <- c()
  
  for (colname in colnames(the_data)) {
     if (colname == "Recurrence.event.s.") next 
     if (colname == "time.to.recurrence") next 
log_rank <- survdiff(Surv(the_data[[the_time]], the_data[[the_event]]) ~ the_data[[colname]])
summary(log_rank)
  if(log_rank$pvalue <= .05) {
    sig_list <- c(sig_list, colname)
    }
  }
  return(sig_list)

  
}
log_rank <- survdiff(Surv(clinData[["time.to.recurrence"]], clinData[["Recurrence.event.s."]]) ~ clinData[["Pathologic.response.N"]])
summary(log_rank)


```

```{r}
sig_list <- log_ranker(car_clinData_valid)
car_clinData_valid_sig <- car_clinData[ , c(sig_list, "time.to.recurrence", "Recurrence.event.s.")]
Kap_creator_significant(car_clinData_valid_sig)

```

## Cox regeression continious variables

### Age at mammo (day)s

```{r}
df <- clinData %>% 
  filter(!is.na(Age.at.mammo..days.))
Surv_obj <-  Surv(time = clinData$time.to.recurrence, event = clinData$Recurrence.event.s.)
cox_demographics <- coxph(Surv_obj ~ as.numeric(Age.at.mammo..days.), 
                   data = clinData)
summary(cox_demographics) #not significa t


```

### size general

```{r}
df <- clinData %>% 
  filter(!is.na(Age.at.mammo..days.))
Surv_obj <-  Surv(time = clinData$time.to.recurrence, event = clinData$Recurrence.event.s.)
cox_demographics <- coxph(Surv_obj ~ as.numeric(size), 
                   data = clinData)
summary(cox_demographics) #not significant


```


