---
title: "Carlos_prelim_anaylsis"
author: "Carlos Haring"
date: "2025-06-18"
output: html_document
---

```{r setup, include=FALSE}
cdc <- read.csv("https://raw.githubusercontent.com/BDSI-CDS-2025/tcia/refs/heads/main/data/clinicalData_clean.csv")
# img <- read.csv("https://raw.githubusercontent.com/BDSI-CDS-2025/tcia/refs/heads/Carlos-Branch/data/imagingFeatures.csv")
```

Load packages that are necessary for running all code below.
```{r}
#' example code for installing packages for first time:
#' install.packages("tidyverse")

library(tidyverse)
library(ggplot2)
library(survival)
library(ggfortify)
library(devtools)
#' install_github("jokergoo/ComplexHeatmap")
library(ComplexHeatmap)
library(umap)
library(gtsummary)
library(summarytools)
library(flextable)
library(forcats)
```


## Reading Clinical Data

```{r}
clinData <- cdc
clinData <- clinData[,-1] # remove first column of row numbers since we have
# patient ID
```

The '<-' is the assignment operator in R. clinData <- read.csv(...) assigns 
the output of read.csv() into a variable/object called clinData. Keyboard 
shortcut on a Mac is option, -. You can also use '='. '<-' is used to
differentiate setting function arguments versus assigning data to a variable.

## Reading Image Features
```{r eval=FALSE, include=FALSE}
imFeatures <- img
imFeatures <- imFeatures[,-1]
```

In general, the feature Citations.docx file describes the features and provides 
citations. The features are grouped into the categories listed below.

Breast and fibroglandular tissue (FGT) volume features: volume and density of 
breast area and fibroglandular area.

Tumor size and morphology: regularity/roundness vs irregularity of tumor
shape and size.

FGT enhancement: measure how much FGT is enhanced when we add contrast.
This is referred to as BPE, background parenchymal enhancement, which may 
confound tumor enhancement/identification.

Tumor enhancement: measure how much tumor is enhanced when we add contrast

Combining tumor and FGT enhancement: measure how tumor and FGT are 
enhanced when adding contrast

FGT enhancement texture: describe the enhancement due to contrast. Looking at
local patterns, does enhancement look gritty or smooth?

Tumor enhancement texture: same as FGT but for tumor area.

Tumor Enhancement Spatial Heterogeneity: measure similarity between tumor
subregions of the tumor.

FGT enhancement variation: global variation of contrast enhancement of FGT

Tumor enhancement variation: global variation of contrast enhancement of tumor

## Task 1: What is in the data?

How many patients are in the dataset?

```{r eval=FALSE, include=FALSE}
nrow(clinData)
#922 patients
```


How many missing values are there among image features?

```{r eval=FALSE, include=FALSE}
sum(is.na(imFeatures))

missingness <- apply(
  imFeatures, 2, function(x){
    sum(is.na(x))
  }
)
sum(missingness == 0) # there are 490 variables with no missingness
sum(missingness != 0) # the remaining 40 variables have some missing values
missingness[missingness != 0]
```

How many missing values are there among clinical features?

```{r eval=FALSE, include=FALSE}
sum(is.na(clinData))

sum(missingness == 0) #490 has no missing values
sum(missigness =! 0) #only one missing value
```


What types of variables are in the clinical dataset? Are they quantitative?

```{r eval=FALSE, include=FALSE}
str(clinData)
clinData %>%
  select(where(is.character)) #36 character column

clinData %>%
  select(where(is.integer)) #59 quantitative

clinData %>%
  select(where(is.factor)) #no factors

clinData %>%
  select(where(is.numeric)) #61 are numeric

ncol(clinData)
```

# Data cleaning 
## changing race
```{r}
clinData <- clinData %>%
  mutate(Race.and.Ethnicity = case_when(Race.and.Ethnicity == 1 ~ "White", Race.and.Ethnicity %in% c(2 ,3, 4, 5, 6, 7) ~ "Other"))

```

##recoding molecular subtype
```{r}
clinData <- clinData %>% 
mutate(Mol.Subtype = case_when(Mol.Subtype == 0 ~ "luminal-like", Mol.Subtype == 1 ~ "ER/PR pos, HER2 pos", Mol.Subtype == 2 ~ "HER2 pos", Mol.Subtype == 3 ~ "triple neg"))
```


##recoding Metastatic at presentation
```{r}
clinData <- clinData %>% 
mutate(Metastatic.at.Presentation..Outside.of.Lymph.Nodes = case_when(Metastatic.at.Presentation..Outside.of.Lymph.Nodes. == 0 ~ "No", Metastatic.at.Presentation..Outside.of.Lymph.Nodes. == 1 ~ "Yes"))
```

##recoding race
```{r}
clinData <- clinData %>%
  mutate(Recurrence.event.s = as.factor(case_when(
      Recurrence.event.s. == 1 ~ "Recurrence",
      Recurrence.event.s. == 0 ~ "No Recurrence")))
```


```{r}
clinData <- clinData[,c(20:56,61:67)]

clinData <- clinData %>%
  mutate(
    Days.to.local.recurrence = as.numeric(na_if(Days.to.local.recurrence..from.the.date.of.diagnosis., "NP")),
    Days.to.distant.recurrence = as.numeric(na_if(Days.to.distant.recurrence.from.the.date.of.diagnosis., "NP")),
    Days.to.last.recurrence.free = as.numeric(na_if(Days.to.last.local.recurrence.free.assessment..from.the.date.of.diagnosis., "NP"))
  ) %>%
  mutate(
    time.to.recurrence = case_when(
      !is.na(Days.to.local.recurrence) & !is.na(Days.to.distant.recurrence) ~
        pmin(Days.to.local.recurrence, Days.to.distant.recurrence, na.rm = TRUE),
      
      !is.na(Days.to.local.recurrence) ~ Days.to.local.recurrence,
      
      !is.na(Days.to.distant.recurrence) ~ Days.to.distant.recurrence,
      
      TRUE ~ Days.to.last.recurrence.free  
    )
  )

clinData <- clinData[,-c(39:47)]
```


```{r,fig.height=20,fig.width=20}
Surv_obj <-  Surv(time = clinData$time.to.recurrence, event = clinData$Recurrence.event.s.)

cox_demographics <- coxph(Surv_obj ~ Date.of.Birth..Days. 
                   + Menopause..at.diagnosis. + Race.and.Ethnicity +
                     Metastatic.at.Presentation..Outside.of.Lymph.Nodes., 
                   data = clinData)
summary(cox_demographics)


KP_race <- survfit(Surv_obj ~ Race.and.Ethnicity, data = clinData)
ggsurvplot(KP_race,
           data = clinData,
           pval = T,
           legend.labs = c("Other", "White"),
           legend.title = "Race and Ethnicity")

KP_meta <- survfit(Surv_obj ~ Metastatic.at.Presentation..Outside.of.Lymph.Nodes., data = clinData)
ggsurvplot(KP_meta,
           data = clinData,
           pval = T,
           legend.labs = c("No Metastasis", "Metastasis Outside Lymph Nodes"), 
           legend.title = "Metastatic Status")

table(clinData$Metastatic.at.Presentation..Outside.of.Lymph.Nodes.)
survdiff(Surv_obj ~ Metastatic.at.Presentation..Outside.of.Lymph.Nodes., data = clinData)
# Race and Metastatic are significant.
```


```{r,fig.height=20,fig.width=20}
clinData <- clinData %>% 
  mutate(Mol.Subtype = as.factor(Mol.Subtype))

cox_tumortype <- coxph(Surv_obj ~ Mol.Subtype, data = clinData)
summary(cox_tumortype)

KP_tumortype <- survfit(Surv_obj ~ Mol.Subtype, data = clinData)

table(clinData$Mol.Subtype)

ggsurvplot(KP_tumortype, 
           data = clinData,
           pval = T,
           legend.labs = c("ER/PR+ HER2+", "HER2-enriched", "Luminal-like", "Triple Negative"),
           legend.title = "Molecular Subtype"
           )

survdiff(Surv_obj ~ Mol.Subtype, data = clinData)

# Triple Negtive are different.
# hazard 2 times

sum(is.na(clinData$Oncotype.score))
# Oncotype DX Recurrence Score too many NAs
```


```{r,fig.height=20,fig.width=20}
clinData <- clinData %>%
  mutate(
    Staging.Tumor.Size....T. = factor(Staging.Tumor.Size....T.,
                     levels = c(1, 2, 3, 4),
                     ordered = TRUE),

    Staging.Nodes...Nx.replaced.by..1..N. = factor(na_if(Staging.Nodes...Nx.replaced.by..1..N., -1),
                     levels = c(0, 1, 2, 3),
                     ordered = TRUE),

    Staging.Metastasis...Mx..replaced.by..1..M. = factor(na_if(Staging.Metastasis...Mx..replaced.by..1..M., -1),
                     levels = c(0, 1),
                     ordered = TRUE)
  )


# cox_stage <- coxph(Surv_obj ~ Staging.Tumor.Size....T. + 
#                      Staging.Nodes...Nx.replaced.by..1..N. +
#                      Staging.Metastasis...Mx..replaced.by..1..M.,
#                    data = clinData)
# summary(cox_stage)


table(clinData$Staging.Tumor.Size....T.)
KP_T <- survfit(Surv_obj ~ Staging.Tumor.Size....T., data = clinData)
ggsurvplot(KP_T, 
           data = clinData,
           pval = T,
           legend.labs = c("T1", "T2", "T3", "T4"),
           legend.title = "Tumor Size"
           )

cox_size <- coxph(Surv_obj ~ Staging.Tumor.Size....T., data = clinData)
summary(cox_size)


table(clinData$Staging.Nodes...Nx.replaced.by..1..N.)
KP_N <- survfit(Surv_obj ~ Staging.Nodes...Nx.replaced.by..1..N., data = clinData)
ggsurvplot(KP_N, 
           data = clinData,
           pval = T,
           legend.labs = c("N0", "N1", "N2", "N3"),
           legend.title = "Nodes (Regional lymph node metastasis status)")

cox_nodes <- coxph(Surv_obj ~ Staging.Nodes...Nx.replaced.by..1..N., data = clinData)
summary(cox_nodes)


table(clinData$Staging.Metastasis...Mx..replaced.by..1..M.)
KP_M <- survfit(Surv_obj ~ Staging.Metastasis...Mx..replaced.by..1..M., data = clinData)
ggsurvplot(KP_M, 
           data = clinData,
           pval = T,
           legend.labs = c("M0", "M1"),
           legend.title = "Metastasis (Distant metastasis status)")

cox_meta <- coxph(Surv_obj ~ Staging.Metastasis...Mx..replaced.by..1..M., data = clinData)
summary(cox_meta)

#all significant
```


```{r,fig.height=20,fig.width=20}
clinData <- clinData %>%
  mutate(
    Nottingham_total_score = as.numeric(TumorGradeT) + as.numeric(TumorGradeN) + as.numeric(TumorGradeM)
  )


KP_nott <- survfit(Surv_obj ~ Nottingham.grade, data = clinData)
ggsurvplot(KP_nott, 
           data = clinData,
           pval = T,
           legend.labs = c("low", "intermediate", "high"),
           legend.title = "Nottingham grade")

cox_nott <- coxph(Surv_obj ~ Nottingham_total_score, data = clinData)
summary(cox_nott)

# It does matter
```


```{r,fig.height=20,fig.width=20}
table(clinData$Histologic.type, clinData$Recurrence.event.s.)


clinData$Histologic.type.new <- fct_collapse(
  factor(clinData$Histologic.type),
  Ductal = c("1"),
  Other = c("0","2", "3", "4", "5", "6", "7", "8", "9", "10")
)

table(clinData$Histologic.type.new, clinData$Recurrence.event.s.)

KP_histologic <- survfit(Surv_obj ~ Histologic.type.new, data = clinData)
ggsurvplot(KP_histologic, 
           data = clinData,
           pval = T,
           legend.labs = c("Other", "Ductal"),
           legend.title = "Histologic type")

cox_histologic <- coxph(Surv_obj ~ Histologic.type.new, data = clinData)
summary(cox_histologic)

# Not sure
# combining some category
```


```{r,fig.height=20,fig.width=20}
clinData <- clinData %>% 
  mutate(Bilateral.Information = as.factor(na_if(Bilateral.Information, "NC")))

table(clinData$Bilateral.Information, clinData$Recurrence.event.s.)

survdiff(Surv_obj ~ Bilateral.Information, data = clinData)

cox_bi <- coxph(Surv_obj ~ Bilateral.Information, data = clinData)
summary(cox_bi)

#Not enough data
```


```{r,fig.height=20,fig.width=20}
clinData <- clinData %>%
  mutate(
    Surgery = as.factor(Surgery),
    Definitive.Surgery.Type = as.factor(na_if(Definitive.Surgery.Type, "NP")),
    Days.to.Surgery..from.the.date.of.diagnosis. = as.numeric(Days.to.Surgery..from.the.date.of.diagnosis.),
    
    Days.to.Surgery = ifelse(
      Days.to.Surgery..from.the.date.of.diagnosis. < 0, 
      NA,  
      Days.to.Surgery..from.the.date.of.diagnosis.
    )
  )

table(clinData$Surgery, clinData$Recurrence.event.s.)
# not enough data for surgery

table(clinData$Definitive.Surgery.Type, clinData$Recurrence.event.s.)

KP_surgery <- survfit(Surv_obj ~ Definitive.Surgery.Type, data = clinData)
ggsurvplot(KP_surgery, 
           data = clinData,
           pval = T,
           legend.labs = c("BCS", "mastectomy"),
           legend.title = "Definitive Surgery Type")

cox_surgery <- coxph(Surv_obj ~ Definitive.Surgery.Type, data = clinData)
summary(cox_surgery)

# 2 times hazard for Mastectomy
```


```{r sensitive analysis}
clinData <- clinData %>%
  mutate(
    time.to.recurrence.new = ifelse(
      !is.na(Days.to.Surgery) & !is.na(time.to.recurrence),
      time.to.recurrence - Days.to.Surgery,
      NA_real_
    )
  )


ggplot(clinData, aes(x = time.to.recurrence, y = time.to.recurrence.new)) +
  geom_point(alpha = 0.8, color = "#A5DEE4", size = 1, shape = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "#D0104C") +
  labs(title = "difference of recurrence time",
       x = "time to recurrence from diagnose",
       y = "time to recurrence from surgery")

```
Definitive.Surgery.Type  
Nottingham.grade 
Staging.Tumor.Size....T. 
Staging.Nodes...Nx.replaced.by..1..N. 
Staging.Metastasis...Mx..replaced.by..1..M. 
Mol.Subtype 
Race
Metastatic



