---
title: "New start"
author: "Carlos"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(naniar)
library(survival)
library(visdat)
library(haven)
library(mice)
library(mitools)
library(purrr)
library(summarytools)
library(gtsummary)
library(randomForestSRC)
library(timeROC)
library(caret)
library(fastDummies)
library(xgboost)
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("survcomp")
library(survcomp)
```

## Data cleaning 

```{r}
data <- read.csv("https://github.com/CarlosZHaring/Recur_Surv/blob/a135a2d47d04819bdf3e05304508bbf34e14f004/cleaned_data.csv")

## Not Include MRI Technical Information.

# First Recurrence 
data <- data %>%
  mutate(
    Days.to.local.recurrence = as.numeric(na_if(Days.to.local.recurrence..from.the.date.of.diagnosis., "NP")),
    Days.to.distant.recurrence = as.numeric(na_if(Days.to.distant.recurrence.from.the.date.of.diagnosis., "NP"))) %>%
  mutate(
    Recurrence.event.s. = case_when(
      Recurrence.event.s. == 1 ~ "Recurrence",
      Recurrence.event.s. == 0 & 
        (!is.na(Days.to.local.recurrence) | 
         !is.na(Days.to.distant.recurrence)) ~ "Recurrence",
      Recurrence.event.s. == 0 ~ "No Recurrence",
      TRUE ~ NA_character_
    ),
    Recurrence.event.s. = factor(Recurrence.event.s., levels = c("No Recurrence", "Recurrence"))
    ) %>%
  mutate(
    Days.to.last.recurrence.free = case_when(
      Recurrence.event.s. == "No Recurrence" ~ as.numeric(na_if(Days.to.last.local.recurrence.free.assessment..from.the.date.of.diagnosis., "NP")),
      TRUE ~ NA_real_
      )
    ) %>%
  mutate(
    dignose.to.recurrence = case_when(
      !is.na(Days.to.local.recurrence) & !is.na(Days.to.distant.recurrence) ~
        pmin(Days.to.local.recurrence, Days.to.distant.recurrence, na.rm = TRUE),
      
      !is.na(Days.to.local.recurrence) ~ Days.to.local.recurrence,
      
      !is.na(Days.to.distant.recurrence) ~ Days.to.distant.recurrence,
      
      TRUE ~ Days.to.last.recurrence.free  
    )
  )




# patient ID
data <- data[,-1] 

data <- data %>%
  mutate(
    mole_type_three = case_when(
     (PR == 1 | ER == 1) & HER2 == 0 ~ "PR+ or ER+",
      HER2 == 1 ~ "HER2+",
      HER2 == 0 & PR == 0 & ER == 0 ~ "Triple Negative"
    ),
    mole_type_three = factor(mole_type_three)
  )



# Race and Metastatic at Presentation (selected)
data<- data %>%
  mutate(Race.and.Ethnicity = as.factor(case_when(Race.and.Ethnicity == 1 ~ "White", 
                                                  Race.and.Ethnicity == 2 ~ "Black",
                                                  Race.and.Ethnicity %in% c(3, 4, 5, 6, 7) ~ "Other")),
         Metastatic.at.Presentation..Outside.of.Lymph.Nodes. = case_when(Metastatic.at.Presentation..Outside.of.Lymph.Nodes. == 0 ~ "No", Metastatic.at.Presentation..Outside.of.Lymph.Nodes. == 1 ~ "Yes"))

# Molecular Type and Tumor Staging (selected)
data <- data %>% mutate(
  Mol.Subtype = as.factor(Mol.Subtype),
  Staging.Tumor.Size....T. = factor(Staging.Tumor.Size....T.,
                     levels = c(1, 2, 3, 4),
                     ordered = TRUE),
  Staging.Nodes...Nx.replaced.by..1..N. = factor(na_if(Staging.Nodes...Nx.replaced.by..1..N., -1),
                     levels = c(0, 1, 2, 3),
                     ordered = TRUE),
  Staging.Metastasis...Mx..replaced.by..1..M. = factor(na_if(Staging.Metastasis...Mx..replaced.by..1..M., -1),
                     levels = c(0, 1),
                     ordered = TRUE),
  Nottingham.total.score = as.numeric(TumorGradeT) + as.numeric(TumorGradeN) + as.numeric(TumorGradeM))

# 72% missing in Oncotype DX (selected)
# Stratify the Oncotype or Imputation
# oncotype_data <- data %>% filter(!is.na(Oncotype.score))

# Collapsed Histologic Type (selected)
data$Histologic.type.collapse <- fct_collapse(
  factor(data$Histologic.type),
  Ductal = c("1"),
  Other = c("0","2", "3", "4", "5", "6", "7", "8", "9", "10")
)

# Bilateral Information (Optional: Detailed Analysis on Subgroup)
data <- data %>% 
  mutate(Bilateral.Information = as.factor(na_if(Bilateral.Information, "NC")))

# MRI Findings (Not Selected Temporarily)
data <- data %>%
  mutate(
    Multicentric.Multifocal = factor(Multicentric.Multifocal, levels = c(0, 1), labels = c("No", "Yes")),
    Contralateral.Breast.Involvement = factor(Contralateral.Breast.Involvement, levels = c(0, 1), labels = c("No", "Yes")),
    Lymphadenopathy.or.Suspicious.Nodes = factor(Lymphadenopathy.or.Suspicious.Nodes, levels = c(0, 1), labels = c("No", "Yes")),
    Skin.Nipple.Invovlement = factor(Skin.Nipple.Invovlement, levels = c(0, 1), labels = c("No", "Yes")),
    Pec.Chest.Involvement = factor(Pec.Chest.Involvement, levels = c(0, 1), labels = c("No", "Yes"))
  )

# Surgery (selected)
data <- data %>%
  mutate(
    # Surgery = as.factor(Surgery),
    Definitive.Surgery.Type = as.factor(na_if(Definitive.Surgery.Type, "NP")),
    Days.to.Surgery..from.the.date.of.diagnosis. = as.numeric(Days.to.Surgery..from.the.date.of.diagnosis.),
    
    Days.to.Surgery = ifelse(
      Days.to.Surgery..from.the.date.of.diagnosis. < 0, 
      NA,  
      Days.to.Surgery..from.the.date.of.diagnosis.
    ),
    surgery.to.recurrence = ifelse(
      !is.na(Days.to.Surgery) & !is.na(dignose.to.recurrence),
      dignose.to.recurrence - Days.to.Surgery,
      NA_real_
    )
  ) 
# %>% filter(!is.na(surgery.to.recurrence))

# Mammography Characteristics and US features (Optional: Detailed Analysis on Subgroup)

# Therapy Type (selected)
data <- data %>%
  mutate(
    Chemotherapy = case_when(
      Neoadjuvant.Chemotherapy == 1 & Adjuvant.Chemotherapy == 1 ~ "Both",
      Neoadjuvant.Chemotherapy == 1 & (Adjuvant.Chemotherapy == 0 | is.na(Adjuvant.Chemotherapy)) ~ "Neoadjuvant Only",
      Adjuvant.Chemotherapy == 1 & (Neoadjuvant.Chemotherapy == 0 | is.na(Neoadjuvant.Chemotherapy)) ~ "Adjuvant Only",
      Neoadjuvant.Chemotherapy == 0 & Adjuvant.Chemotherapy == 0 ~ "None",
      TRUE ~ NA_character_
    ) %>% as.factor(),

    Radiationtherapy = case_when(
      Neoadjuvant.Radiation.Therapy == 1 & Adjuvant.Radiation.Therapy == 1 ~ "Both",
      Neoadjuvant.Radiation.Therapy == 1 & (Adjuvant.Radiation.Therapy == 0 | is.na(Adjuvant.Radiation.Therapy)) ~ "Neoadjuvant Only",
      Adjuvant.Radiation.Therapy == 1 & (Neoadjuvant.Radiation.Therapy == 0 | is.na(Neoadjuvant.Radiation.Therapy)) ~ "Adjuvant Only",
      Neoadjuvant.Radiation.Therapy == 0 & Adjuvant.Radiation.Therapy == 0 ~ "None",
      TRUE ~ NA_character_
    ) %>% as.factor(),

    Endocrinetherapy = case_when(
      Neoadjuvant.Endocrine.Therapy.Medications == 1 & Adjuvant.Endocrine.Therapy.Medications == 1 ~ "Both",
      Neoadjuvant.Endocrine.Therapy.Medications == 1 & (Adjuvant.Endocrine.Therapy.Medications == 0 | is.na(Adjuvant.Endocrine.Therapy.Medications)) ~ "Neoadjuvant Only",
      Adjuvant.Endocrine.Therapy.Medications == 1 & (Neoadjuvant.Endocrine.Therapy.Medications == 0 | is.na(Neoadjuvant.Endocrine.Therapy.Medications)) ~ "Adjuvant Only",
      Neoadjuvant.Endocrine.Therapy.Medications == 0 & Adjuvant.Endocrine.Therapy.Medications == 0 ~ "None",
      TRUE ~ NA_character_
    ) %>% as.factor(),
    
    Antiher2therapy = case_when(
      Neoadjuvant.Anti.Her2.Neu.Therapy == 1 & Adjuvant.Anti.Her2.Neu.Therapy == 1 ~ "Both",
      Neoadjuvant.Anti.Her2.Neu.Therapy == 1 & (Adjuvant.Anti.Her2.Neu.Therapy == 0 | is.na(Adjuvant.Anti.Her2.Neu.Therapy)) ~ "Neoadjuvant Only",
      Adjuvant.Anti.Her2.Neu.Therapy == 1 & (Neoadjuvant.Anti.Her2.Neu.Therapy == 0 | is.na(Neoadjuvant.Anti.Her2.Neu.Therapy)) ~ "Adjuvant Only",
      Neoadjuvant.Anti.Her2.Neu.Therapy == 0 & Adjuvant.Anti.Her2.Neu.Therapy == 0 ~ "None",
      TRUE ~ NA_character_
    ) %>% as.factor()
    
  # Antiher2therapy_b = case_when(Neoadjuvant.Anti.Her2.Neu.Therapy == 1 &  Adjuvant.Anti.Her2.Neu.Therapy == 1 ~ "Her2 Therapy",
  #   Neoadjuvant.Anti.Her2.Neu.Therapy == 1 ~ "Her2 Therapy", 
  #   Adjuvant.Anti.Her2.Neu.Therapy == 1 ~ "Her2 Therapy", 
  #   Neoadjuvant.Anti.Her2.Neu.Therapy == 0 ~ "No Her2 Therapy", 
  #   Adjuvant.Anti.Her2.Neu.Therapy == 0 ~ "No Her2 Therapy",
  #   TRUE ~ NA_character_
  # ) %>% as.factor(),
  # 
  # Endocrinetherapy_b = case_when(Neoadjuvant.Endocrine.Therapy.Medications == 1 & Adjuvant.Endocrine.Therapy.Medications == 1 ~ "Endocrine Therapy", 
  #   Neoadjuvant.Endocrine.Therapy.Medications == 1 ~ "Endocrine Therapy", 
  #   Adjuvant.Endocrine.Therapy.Medications == 1 ~ "Endocrine Therapy", 
  #   Neoadjuvant.Endocrine.Therapy.Medications == 0 ~ " No Endocrine Therapy", 
  #   Adjuvant.Endocrine.Therapy.Medications == 0 ~ "No Endocrine Therapy",
  #   TRUE ~ NA_character_
  #   ) %>% as.factor(),
  # 
  # Chemotherapy_b = case_when(Adjuvant.Chemotherapy == 1 & Neoadjuvant.Chemotherapy == 1 ~ "Chemotherapy",
  #   Adjuvant.Chemotherapy == 1 ~ "Chemotherapy", 
  #   Neoadjuvant.Chemotherapy == 1 ~ "Chemotherapy", 
  #   Adjuvant.Chemotherapy == 0 ~ "No Chemotherapy", 
  #   Neoadjuvant.Chemotherapy == 0~ "No Chemotherapy",
  #   TRUE ~ NA_character_
  #   ) %>% as.factor(),
  # 
  # Radiationtherapy_b = case_when(Adjuvant.Chemotherapy == 1 & Neoadjuvant.Chemotherapy == 1 ~ "Chemotherapy",
  #   Adjuvant.Radiation.Therapy == 1 ~ "Radiation Therapy", 
  #   Neoadjuvant.Endocrine.Therapy.Medications == 1 ~ "Radiation Therapy", 
  #   Adjuvant.Radiation.Therapy == 0 ~ "No Radiation Therapy", 
  #   Neoadjuvant.Endocrine.Therapy.Medications == 0 ~"No Radiation Therapy",
  #   TRUE ~ NA_character_
  #   ) %>% as.factor(),
  )
  
  



# Ovarian: Related to Endocrine Therapy (selected)
data <- data %>%
  mutate(
    Known.Ovarian.Status = factor(Known.Ovarian.Status,
                                   levels = c(0, 1),
                                   labels = c("No", "Yes")),
    
    Number.of.Ovaries.In.Situ.. = na_if(Number.of.Ovaries.In.Situ.., "NP"),
    Number.of.Ovaries.In.Situ.. = factor(as.numeric(Number.of.Ovaries.In.Situ..),
                                         levels = c(0, 1, 2),
                                         labels = c("0 ovaries", "1 ovary", "2 ovaries"),
                                         ordered = TRUE),
    
    Therapeutic.or.Prophylactic.Oophorectomy.as.part.of.Endocrine.Therapy =
      factor(Therapeutic.or.Prophylactic.Oophorectomy.as.part.of.Endocrine.Therapy,
             levels = c(0, 1),
             labels = c("No", "Yes"))
  )

# Neoadjuvant Therapy or not (selected)
data <- data %>%
  mutate(
    Received.Neoadjuvant.Therapy.or.Not = if_else(
      Chemotherapy %in% c("Neoadjuvant Only", "Both") |
      Radiationtherapy %in% c("Neoadjuvant Only", "Both") |
      Endocrinetherapy %in% c("Neoadjuvant Only", "Both") |
      Antiher2therapy %in% c("Neoadjuvant Only", "Both"),
      1, 2
    )
  ) %>%
  mutate(Received.Neoadjuvant.Therapy.or.Not = factor(case_when(
    Received.Neoadjuvant.Therapy.or.Not == 1 ~ "Yes",
    Received.Neoadjuvant.Therapy.or.Not == 2 ~ "No",
    TRUE ~ NA_character_
  )))


## All the Variable Below Must be Analysis Using Subgroup (all below selected)

# Tumor Response
data <- data %>%
  mutate(Clinical.Response..Evaluated.Through.Imaging = case_when(
    Clinical.Response..Evaluated.Through.Imaging == 1 ~ "Complete",
    Clinical.Response..Evaluated.Through.Imaging == 2 ~ "Not Complete",
    Clinical.Response..Evaluated.Through.Imaging == 3 ~ NA_character_,  
    TRUE ~ NA_character_
  ),
  Clinical.Response..Evaluated.Through.Imaging = factor(Clinical.Response..Evaluated.Through.Imaging, levels = c("Not Complete", "Complete"))) %>%
  mutate(Pathologic.Response.to.Neoadjuvant.Therapy = case_when(
    Pathologic.Response.to.Neoadjuvant.Therapy == 1 ~ "Complete",
    Pathologic.Response.to.Neoadjuvant.Therapy == 2 ~ "Not Complete",
    Pathologic.Response.to.Neoadjuvant.Therapy == 3 ~ "DCIS Only",
    Pathologic.Response.to.Neoadjuvant.Therapy == 4 ~ "LCIS Only",
    Pathologic.Response.to.Neoadjuvant.Therapy == 5 ~ NA_character_,
    TRUE ~ NA_character_
  ),
  Pathologic.Response.to.Neoadjuvant.Therapy = factor(Pathologic.Response.to.Neoadjuvant.Therapy,
                               levels = c("Not Complete", "DCIS Only", "LCIS Only", "Complete")))

# Pathologic Stage 
data <- data %>%
  mutate(
    Pathologic.response.to.Neoadjuvant.therapy..Pathologic.stage..T..following.neoadjuvant.therapy = na_if(Pathologic.response.to.Neoadjuvant.therapy..Pathologic.stage..T..following.neoadjuvant.therapy, -1),
    Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..N..following.neoadjuvant.therapy = na_if(Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..N..following.neoadjuvant.therapy, -1)
  ) %>%
  mutate(
    Pathologic.response.to.Neoadjuvant.therapy..Pathologic.stage..T..following.neoadjuvant.therapy = factor(Pathologic.response.to.Neoadjuvant.therapy..Pathologic.stage..T..following.neoadjuvant.therapy, levels = c(0, 1, 2, 3, 4, 5), ordered = TRUE),
    Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..N..following.neoadjuvant.therapy = factor(Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..N..following.neoadjuvant.therapy, levels = c(0, 1, 2, 3), ordered = TRUE)
  ) %>%
  mutate(
    Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..M..following.neoadjuvant.therapy = case_when(
      Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..M..following.neoadjuvant.therapy == -1 ~ "MX",
      Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..M..following.neoadjuvant.therapy == 0 ~ "M0",
      Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..M..following.neoadjuvant.therapy == 1 ~ "M1",
      TRUE ~ NA_character_
    ),
    Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..M..following.neoadjuvant.therapy = factor(Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..M..following.neoadjuvant.therapy,
                                   levels = c("M0", "M1", "MX"),
                                   ordered = FALSE)
  )

# Near Complete Response
data <- data %>%
  mutate(
    Overall.Near.complete.Response...Stricter.Definition = case_when(
      Overall.Near.complete.Response...Stricter.Definition == 3 ~ NA_real_,
      TRUE ~ Overall.Near.complete.Response...Stricter.Definition
    ),
    Overall.Near.complete.Response...Looser.Definition = case_when(
      Overall.Near.complete.Response...Looser.Definition == 3 ~ NA_real_,
      TRUE ~ Overall.Near.complete.Response...Looser.Definition
    ),
    Near.complete.Response..Graded.Measure. = case_when(
      Near.complete.Response..Graded.Measure. == 19 ~ NA_real_,
      TRUE ~ Near.complete.Response..Graded.Measure.
    ),
    
    Overall.Near.complete.Response...Stricter.Definition = factor(
      Overall.Near.complete.Response...Stricter.Definition,
      levels = c(0, 1, 2),
      labels = c("Not complete", "Complete", "Near-complete")
    ),
    
    Overall.Near.complete.Response...Looser.Definition = factor(
      Overall.Near.complete.Response...Looser.Definition,
      levels = c(0, 1, 2),
      labels = c("Not complete", "Complete", "Near-complete")
    ),
    
    Near.complete.Response..Graded.Measure. = factor(
      Near.complete.Response..Graded.Measure.
    ),
    Metastatic.at.Presentation..Outside.of.Lymph.Nodes. = factor(
    Metastatic.at.Presentation..Outside.of.Lymph.Nodes.)
  )

names(data)
```


```{r}
cleaned_data <- data %>% select(
  "Race.and.Ethnicity",
  "Metastatic.at.Presentation..Outside.of.Lymph.Nodes.",
  "Mol.Subtype",
  "Staging.Tumor.Size....T.",
  "Staging.Nodes...Nx.replaced.by..1..N.",
  "Staging.Metastasis...Mx..replaced.by..1..M.",
  "Nottingham.total.score",
  "Oncotype.score",
  "Histologic.type.collapse",
  "Surgery",
  "Definitive.Surgery.Type",
  "Chemotherapy",
  "Radiationtherapy",
  "Endocrinetherapy",
  "Antiher2therapy",
  #"Chemotherapy_b",
  #"Radiationtherapy_b",
  #"Endocrinetherapy_b",
  #"Antiher2therapy_b",
  "Known.Ovarian.Status",
  "Number.of.Ovaries.In.Situ..",
  "Therapeutic.or.Prophylactic.Oophorectomy.as.part.of.Endocrine.Therapy",
  "Received.Neoadjuvant.Therapy.or.Not",
  "Clinical.Response..Evaluated.Through.Imaging",
  "Pathologic.Response.to.Neoadjuvant.Therapy",
  "Pathologic.response.to.Neoadjuvant.therapy..Pathologic.stage..T..following.neoadjuvant.therapy",
  "Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..N..following.neoadjuvant.therapy",
  "Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..M..following.neoadjuvant.therapy",
  "Overall.Near.complete.Response...Stricter.Definition",
  "Overall.Near.complete.Response...Looser.Definition",
  "Near.complete.Response..Graded.Measure.",
  "Days.to.Surgery",
  "dignose.to.recurrence",
  "Recurrence.event.s.",
  "surgery.to.recurrence",
  "Age.at.mammo..days.",
  "mole_type_three"
  
) %>% 
  rename(
    Race = Race.and.Ethnicity,
    Metastatic = Metastatic.at.Presentation..Outside.of.Lymph.Nodes.,
    Mole_type = Mol.Subtype,
    T_stage = Staging.Tumor.Size....T.,
    N_stage = Staging.Nodes...Nx.replaced.by..1..N.,
    M_stage = Staging.Metastasis...Mx..replaced.by..1..M.,
    Nottingham_score = Nottingham.total.score,
    Oncotype_score = Oncotype.score,
    Histology = Histologic.type.collapse,
    Surgery = Surgery,
    Surgery_type = Definitive.Surgery.Type,
    Chemo = Chemotherapy,
    Radiation = Radiationtherapy,
    Endocrine = Endocrinetherapy,
    Antiher2 = Antiher2therapy,
    Ovarian_status = Known.Ovarian.Status,
    Ovary_count = Number.of.Ovaries.In.Situ..,
    Oophorectomy = Therapeutic.or.Prophylactic.Oophorectomy.as.part.of.Endocrine.Therapy,
    Neoadjuvant = Received.Neoadjuvant.Therapy.or.Not,
    Clinical_response = Clinical.Response..Evaluated.Through.Imaging,
    Pathologic_response = Pathologic.Response.to.Neoadjuvant.Therapy,
    Patho_T_stage = Pathologic.response.to.Neoadjuvant.therapy..Pathologic.stage..T..following.neoadjuvant.therapy,
    Patho_N_stage = Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..N..following.neoadjuvant.therapy,
    Patho_M_stage = Pathologic.response.to.Neoadjuvant.therapy...Pathologic.stage..M..following.neoadjuvant.therapy,
    Near_response_strict = Overall.Near.complete.Response...Stricter.Definition,
    Near_response_loose = Overall.Near.complete.Response...Looser.Definition,
    Near_response_grade = Near.complete.Response..Graded.Measure.,
    Time_to_surgery = Days.to.Surgery,
    Dignose_to_recurrence = dignose.to.recurrence,
    Recurrence = Recurrence.event.s.,
    Time_to_event = surgery.to.recurrence,
    Age_at_mammo = Age.at.mammo..days.,
    mole_type_three = mole_type_three
  )
```

# more cleaning and filtering 
## visualize neoadjuvent data

```{r}
data <- read.csv("https://github.com/CarlosZHaring/Recur_Surv/blob/a135a2d47d04819bdf3e05304508bbf34e14f004/cleaned_data.csv")
clean<- cleaned_data %>%
  filter(Surgery == 1) %>%
  filter(!is.na(Surgery)) %>%
  filter(!is.na(Neoadjuvant)) %>%
  filter(`Neoadjuvant` == "Yes") %>%
  filter(!is.na(Time_to_event))

  clean %>%
    group_by(Recurrence) %>%
    count()

clean$Neoadjuvant

library(naniar)
library(ggplot2)

p <- vis_miss(clean, sort_miss = TRUE) +
  labs(title = "Missingness among those of neoadjuvant therapy") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    plot.margin = margin(t = 10, r = 40, b = 10, l = 10)  # add space on the right
  )

ggsave("missingness_plot.png", plot = p, width = 10, height = 6, units = "in", dpi = 300)
p
```


## mutation of data
```{r}

clean <- clean %>%
  mutate(Patho_N_binary = case_when(
    Patho_N_stage %in% c("0", "1") ~ "No/small lymph involvment",
    Patho_N_stage %in% c("2", "3") ~ "medium/heavy lymph involvment",
    TRUE ~ NA_character_
  )) %>%
  
  mutate(Patho_N_binary = factor(Patho_N_binary)) %>%
  
  mutate(Patho_T_binary = case_when(
    Patho_T_stage %in% c("1", "2", "3", "4") ~ "Present",
    Patho_T_stage %in% c("0", "5") ~ "Absent_or_Pre-invasive",
    TRUE ~ NA_character_
  )) %>%
  
  mutate(Patho_T_binary = factor(Patho_T_binary)) %>%
  
  mutate(Patho_M_Binary = case_when(
    Patho_M_stage %in% c("MX", "M0") ~ "No Metastasis",
    Patho_M_stage %in% c("M1") ~ "Metastasis",
    TRUE ~ NA_character_
  )) %>%
  
  mutate(Patho_M_Binary = factor(Patho_M_Binary)) %>%
  mutate(Nottingham_score_binary = case_when(
    Nottingham_score %in% c("3", "4", "5", "6", "7") ~ "Low score",
    Nottingham_score %in% c("8", "9") ~ "High Score",
    TRUE ~ NA_character_
  )) %>%
  
  mutate(Nottingham_score_binary = factor(Nottingham_score_binary)) %>%
  
  mutate(Race_binary = case_when(
    Race %in% "White" ~ "White",
    TRUE ~ "Other",
    TRUE ~ NA_character_
  )) %>%
  
  mutate(Race_binary = factor(Race_binary)) %>%
  
  mutate(T_stage_binary = case_when(
    T_stage %in% c("1", "2") ~ "Small",
    T_stage %in% c("3", "4") ~ "Large",
    TRUE ~ NA_character_
  )) %>%
  
  mutate(T_stage_binary = factor(T_stage_binary)) %>%
  
  mutate(Antiher2_treatment = case_when(
    Antiher2 == "None" ~ "No Treatment",
    Antiher2 %in% c("Adjuvant Only", "Both", "Neoadjuvant Only") ~ "Treatment",
    TRUE ~ NA_character_
  )) %>%
  
  mutate(Antiher2_treatment = factor(Antiher2_treatment)) %>%
  
  mutate(Mole_type_tri = case_when(
    Mole_type %in% c("1", "2") ~ "HER+",
    Mole_type == "0" ~ "Luminal-like",
    Mole_type == "3" ~ "Triple Negative",
    TRUE ~ NA_character_
  )) %>%
  
  mutate(Mole_type_tri = factor(Mole_type_tri)) %>%
  
  mutate(Chemo_adjuvent = case_when(
    Chemo %in% c("Adjuvant Only", "Both") ~ "Adjuvent Treatment",
    Chemo %in% c("Neoadjuvant Only" ,"None") ~ "No Adjuvent Treatment",
    TRUE ~ NA_character_)) %>%
  
  mutate(Chemo_adjuvent =factor(Chemo_adjuvent)) %>%
  
  mutate(Near_response_strict_binary = case_when(
    Near_response_strict %in% c("Not complete", "Near-complete") ~ "Not Complete",
    Near_response_strict == "Complete" ~ "Complete",
    TRUE ~ NA_character_
  )) %>%
  
  mutate(Near_response_strict_binary = factor(Near_response_strict_binary)) %>%
  
  mutate(Radiation_adjuvent = case_when(
    Radiation == "Adjuvant Only" ~ "Adjuvent Treatment",
   Radiation  %in% c("Neoadjuvant", "None") ~ "No Adjuvent Treatment")) %>%
  
  mutate(Radiation_adjuvent =factor(Radiation_adjuvent))

```



## filter

```{r}
clean <- clean %>% #histology was removed since it was not converging 
  select(c(Histology,
    Patho_N_stage, Patho_T_stage, Patho_M_stage, Nottingham_score, Endocrine, Race, Pathologic_response, Near_response_strict, Near_response_loose, Near_response_grade, T_stage, Antiher2, N_stage, Metastatic, Mole_type, Surgery_type, Chemo, Radiation, Ovarian_status, Oophorectomy, Neoadjuvant, Time_to_surgery, Dignose_to_recurrence, Recurrence, Time_to_event, Surgery, Radiation_adjuvent, Chemo_adjuvent, Mole_type_tri, Antiher2_treatment, T_stage_binary, Race_binary, Nottingham_score_binary, Patho_M_Binary, Patho_T_binary, Patho_N_binary, Near_response_strict_binary, Clinical_response, Age_at_mammo, mole_type_three))

vis_miss(clean, sort_miss = TRUE)

# Convert problematic ordered factors to regular factors or numeric
clean <- clean %>%
  mutate(
    # Convert ordered factors to regular factors to avoid polr issues
    T_stage = as.factor(as.character(T_stage)),
    N_stage = as.factor(as.character(N_stage)),
    Patho_T_stage = as.factor(as.character(Patho_T_stage)),
    Patho_N_stage = as.factor(as.character(Patho_N_stage)),
    Ovarian_status = as.factor(as.character(Ovarian_status))
  ) %>%
  # Drop unused factor levels
  mutate(across(where(is.factor), ~fct_drop(.))) %>%
  mutate(Recurrence = recode(Recurrence, "No Recurrence" = 0, "Recurrence" = 1)) %>%
  mutate(Recurrence = as.numeric(Recurrence))

clean$Recurrence

```



# Model Trails (do not run these chunks all at once, run each chunk individaully to see model outputs)

## original variable models

### first model
```{r}
set.seed(123)
folds <- createFolds(clean$Recurrence, k = 4, returnTrain = TRUE)
cindex_vals <- numeric(length(folds))
vimp_list <- vector("list", length(folds))
for (i in seq_along(folds)) {
  train_data <- clean[folds[[1]], ]
  test_data  <- clean[-folds[[1]], ]
  
  mod <- rfsrc(
    Surv(time = Time_to_event, event = Recurrence) ~ 
      Race + Mole_type + Nottingham_score + Surgery + Surgery_type + 
      Chemo + Antiher2 + Endocrine + Oophorectomy + Ovarian_status + 
      Neoadjuvant + Pathologic_response + Patho_N_stage + 
      Patho_T_stage + Patho_M_stage,
    data = train_data,
    ntree = 1000,
    na.action = "na.impute"
  )
 vimp_vals <- vimp(mod)$importance
  vimp_list[[i]] <- vimp_vals
  # ave prediction result
pred <- predict(mod, newdata = test_data, na.action = "na.impute")
  # Extract risk scores
  risk_scores <- pred$predicted
  nrow(test_data)
length(risk_scores)
  # Create survival object for test data
  surv_obj <- Surv(test_data$Time_to_event, test_data$Recurrence)

  # Compute concordance index
  cindex_vals[i] <- survConcordance(surv_obj ~ risk_scores)$concordance
}
# Average C-index across folds
mean_cindex <- mean(cindex_vals)
print(mean_cindex)
vimp_df <- do.call(cbind, vimp_list)
vimp_summary <- data.frame(
  Variable = rownames(vimp_df),
  Mean = rowMeans(vimp_df, na.rm = TRUE),
  SD = apply(vimp_df, 1, sd, na.rm = TRUE)
)

vimp_summary %>%
  arrange(Mean)
```

### second model
```{r}
set.seed(123)
folds <- createFolds(clean$Recurrence, k = 4, returnTrain = TRUE)
cindex_vals <- numeric(length(folds))
vimp_list <- vector("list", length(folds))
for (i in seq_along(folds)) {
  train_data <- clean[folds[[1]], ]
  test_data  <- clean[-folds[[1]], ]
  
  mod <- rfsrc(
    Surv(time = Time_to_event, event = Recurrence) ~ 
      Mole_type + Nottingham_score + Surgery_type + 
      Chemo + Antiher2 + Endocrine + Oophorectomy + 
       Pathologic_response + Patho_N_stage + 
      Patho_T_stage,
    data = train_data,
    ntree = 1000,
    na.action = "na.impute",
    
  )
 vimp_vals <- vimp(mod)$importance
  vimp_list[[i]] <- vimp_vals
  # ave prediction result
pred <- predict(mod, newdata = test_data, na.action = "na.impute")
  # Extract risk scores
  risk_scores <- pred$predicted
  nrow(test_data)
length(risk_scores)
  # Create survival object for test data
  surv_obj <- Surv(test_data$Time_to_event, test_data$Recurrence)

  # Compute concordance index
  cindex_vals[i] <- survConcordance(surv_obj ~ risk_scores)$concordance
}
# Average C-index across folds
mean_cindex <- mean(cindex_vals)
print(mean_cindex)
vimp_df <- do.call(cbind, vimp_list)
vimp_summary <- data.frame(
  Variable = rownames(vimp_df),
  Mean = rowMeans(vimp_df, na.rm = TRUE),
  SD = apply(vimp_df, 1, sd, na.rm = TRUE)
)

vimp_summary %>%
  arrange(Mean)
```

### third model
```{r}
set.seed(123)
folds <- createFolds(clean$Recurrence, k = 4, returnTrain = TRUE)
cindex_vals <- numeric(length(folds))
vimp_list <- vector("list", length(folds))
for (i in seq_along(folds)) {
  train_data <- clean[folds[[1]], ]
  test_data  <- clean[-folds[[1]], ]
  
  mod <- rfsrc(
    Surv(time = Time_to_event, event = Recurrence) ~ 
      Mole_type + Nottingham_score + 
      Chemo + Antiher2 + 
       Pathologic_response + Patho_N_stage + 
      Patho_T_stage,
    data = train_data,
    ntree = 1000,
    na.action = "na.impute",
    
  )
 vimp_vals <- vimp(mod)$importance
  vimp_list[[i]] <- vimp_vals
  # Save prediction result
pred <- predict(mod, newdata = test_data, na.action = "na.impute")
  # Extract risk scores
  risk_scores <- pred$predicted
  nrow(test_data)
length(risk_scores)
  # Create survival object for test data
  surv_obj <- Surv(test_data$Time_to_event, test_data$Recurrence)

  # Compute concordance index
  cindex_vals[i] <- survConcordance(surv_obj ~ risk_scores)$concordance
}
# Average C-index across folds
mean_cindex <- mean(cindex_vals)
print(mean_cindex)
vimp_df <- do.call(cbind, vimp_list)
vimp_summary <- data.frame(
  Variable = rownames(vimp_df),
  Mean = rowMeans(vimp_df, na.rm = TRUE),
  SD = apply(vimp_df, 1, sd, na.rm = TRUE)
)

vimp_summary %>%
  arrange(Mean)
```


## dichotomous variable models

### model one
```{r}
set.seed(123)
folds <- createFolds(clean$Recurrence, k = 4, returnTrain = TRUE)
cindex_vals <- numeric(length(folds))
vimp_list <- vector("list", length(folds))
for (i in seq_along(folds)) {
  train_data <- clean[folds[[1]], ]
  test_data  <- clean[-folds[[1]], ]
  
  mod <- rfsrc(
    Surv(time = Time_to_event, event = Recurrence) ~ 
      Race_binary + Mole_type_tri + Nottingham_score_binary + Surgery + Surgery_type + 
      Chemo_adjuvent + Antiher2_treatment + Endocrine + Oophorectomy + Ovarian_status + 
      Neoadjuvant + Pathologic_response + Patho_N_binary + 
      Patho_T_binary + Patho_M_Binary + Near_response_strict_binary,
    data = train_data,
    ntree = 1000,
    na.action = "na.impute"
  )
 vimp_vals <- vimp(mod)$importance
  vimp_list[[i]] <- vimp_vals
  # ✅ Save prediction result
pred <- predict(mod, newdata = test_data, na.action = "na.impute")
  # Extract risk scores
  risk_scores <- pred$predicted
  nrow(test_data)
length(risk_scores)
  # Create survival object for test data
  surv_obj <- Surv(test_data$Time_to_event, test_data$Recurrence)

  # Compute concordance index
  cindex_vals[i] <- survConcordance(surv_obj ~ risk_scores)$concordance
}
# Average C-index across folds
mean_cindex <- mean(cindex_vals)
print(mean_cindex)
vimp_df <- do.call(cbind, vimp_list)
vimp_summary <- data.frame(
  Variable = rownames(vimp_df),
  Mean = rowMeans(vimp_df, na.rm = TRUE),
  SD = apply(vimp_df, 1, sd, na.rm = TRUE)
)

vimp_summary %>%
  arrange(Mean)
```

### model two 
```{r}

set.seed(123)
folds <- createFolds(clean$Recurrence, k = 4, returnTrain = TRUE)
cindex_vals <- numeric(length(folds))
vimp_list <- vector("list", length(folds))
for (i in seq_along(folds)) {
  train_data <- clean[folds[[1]], ]
  test_data  <- clean[-folds[[1]], ]
  
  mod <- rfsrc(
    Surv(time = Time_to_event, event = Recurrence) ~ 
      Race_binary + Mole_type_tri + Nottingham_score_binary + 
      Chemo_adjuvent + Antiher2_treatment + Endocrine + Oophorectomy + Pathologic_response + Patho_N_binary + 
      Patho_T_binary + Near_response_strict_binary,
    data = train_data,
    ntree = 1000,
    na.action = "na.impute"
  )
 vimp_vals <- vimp(mod)$importance
  vimp_list[[i]] <- vimp_vals
  # ave prediction result
pred <- predict(mod, newdata = test_data, na.action = "na.impute")
  # Extract risk scores
  risk_scores <- pred$predicted
  nrow(test_data)
length(risk_scores)
  # Create survival object for test data
  surv_obj <- Surv(test_data$Time_to_event, test_data$Recurrence)

  # Compute concordance index
  cindex_vals[i] <- survConcordance(surv_obj ~ risk_scores)$concordance
}
# Average C-index across folds
mean_cindex <- mean(cindex_vals)
print(mean_cindex)
vimp_df <- do.call(cbind, vimp_list)
vimp_summary <- data.frame(
  Variable = rownames(vimp_df),
  Mean = rowMeans(vimp_df, na.rm = TRUE),
  SD = apply(vimp_df, 1, sd, na.rm = TRUE)
)

vimp_summary %>%
  arrange(Mean)
```


### model three 
```{r}

set.seed(123)
folds <- createFolds(clean$Recurrence, k = 4, returnTrain = TRUE)
cindex_vals <- numeric(length(folds))
vimp_list <- vector("list", length(folds))
for (i in seq_along(folds)) {
  train_data <- clean[folds[[1]], ]
  test_data  <- clean[-folds[[1]], ]
  
  mod <- rfsrc(
    Surv(time = Time_to_event, event = Recurrence) ~ Mole_type_tri + Nottingham_score_binary + 
      Chemo_adjuvent + Antiher2_treatment + Endocrine + Oophorectomy + Pathologic_response + Patho_N_binary + Near_response_strict_binary,
    data = train_data,
    ntree = 1000,
    na.action = "na.impute"
  )
 vimp_vals <- vimp(mod)$importance
  vimp_list[[i]] <- vimp_vals
  # ave prediction result
pred <- predict(mod, newdata = test_data, na.action = "na.impute")
  # Extract risk scores
  risk_scores <- pred$predicted
  nrow(test_data)
length(risk_scores)
  # Create survival object for test data
  surv_obj <- Surv(test_data$Time_to_event, test_data$Recurrence)

  # Compute concordance index
  cindex_vals[i] <- survConcordance(surv_obj ~ risk_scores)$concordance
}
# Average C-index across folds
mean_cindex <- mean(cindex_vals)
print(mean_cindex)
vimp_df <- do.call(cbind, vimp_list)
vimp_summary <- data.frame(
  Variable = rownames(vimp_df),
  Mean = rowMeans(vimp_df, na.rm = TRUE),
  SD = apply(vimp_df, 1, sd, na.rm = TRUE)
)

vimp_summary %>%
  arrange(Mean)
```

### model 4
```{r}
set.seed(123)
folds <- createFolds(clean$Recurrence, k = 4, returnTrain = TRUE)
cindex_vals <- numeric(length(folds))
vimp_list <- vector("list", length(folds))
for (i in seq_along(folds)) {
  train_data <- clean[folds[[1]], ]
  test_data  <- clean[-folds[[1]], ]
  
  mod <- rfsrc(
    Surv(time = Time_to_event, event = Recurrence) ~ Mole_type_tri + Nottingham_score_binary + Antiher2_treatment + Endocrine + Oophorectomy + Pathologic_response + Patho_N_binary + Near_response_strict_binary,
    data = train_data,
    ntree = 1000,
    na.action = "na.impute"
  )
 vimp_vals <- vimp(mod)$importance
  vimp_list[[i]] <- vimp_vals
  # ave prediction result
pred <- predict(mod, newdata = test_data, na.action = "na.impute")
  # Extract risk scores
  risk_scores <- pred$predicted
  nrow(test_data)
length(risk_scores)
  # Create survival object for test data
  surv_obj <- Surv(test_data$Time_to_event, test_data$Recurrence)

  # Compute concordance index
  cindex_vals[i] <- survConcordance(surv_obj ~ risk_scores)$concordance
}
# Average C-index across folds
mean_cindex <- mean(cindex_vals)
print(mean_cindex)
vimp_df <- do.call(cbind, vimp_list)
vimp_summary <- data.frame(
  Variable = rownames(vimp_df),
  Mean = rowMeans(vimp_df, na.rm = TRUE),
  SD = apply(vimp_df, 1, sd, na.rm = TRUE)
)

vimp_summary %>%
  arrange(Mean)
```

### model 5
```{r}
seed <- sample(seq(from = 1, to = 10000), size = 30, replace = FALSE)
saveRDS(seeds, "seed.txt")
list = NULL
for (s in seeds) {
  set.seed(s)
  folds <- createFolds(clean$Recurrence, k = 4, returnTrain = TRUE)
  cindex_vals <- numeric(length(folds))
  vimp_list <- vector("list", length(folds))
  
  for (j in seq_along(folds)) {
    train_data <- clean[folds[[j]], ]
    test_data  <- clean[-folds[[j]], ]
    
    mod <- rfsrc(
     Surv(Time_to_event, Recurrence) ~ Mole_type_tri + T_stage_binary +
  N_stage + Antiher2_treatment + Endocrine + Oophorectomy +
  Pathologic_response + Patho_N_binary + Near_response_strict_binary,
      data = train_data,
      ntree = 1000,
      na.action = "na.impute"
    )
    
    vimp_vals <- vimp(mod)$importance
    vimp_list[[j]] <- vimp_vals
    
    pred <- predict(mod, newdata = test_data, na.action = "na.impute")
    risk_scores <- pred$predicted
    surv_obj <- Surv(test_data$Time_to_event, test_data$Recurrence)
    cindex_vals[j] <- survConcordance(surv_obj ~ risk_scores)$concordance
  }
  
  # Store average c-index and VIMP summary
  mean_cindex <- mean(cindex_vals)
  print(mean_cindex)
  
  vimp_df <- do.call(cbind, vimp_list)
  vimp_summary <- data.frame(
    Variable = rownames(vimp_df),
    Mean = rowMeans(vimp_df, na.rm = TRUE),
    SD = apply(vimp_df, 1, sd, na.rm = TRUE)
  )
  
  vimp_summary <- vimp_summary %>% arrange(Mean)
  list[[as.character(s)]] <- list(cindex = mean_cindex, vimp = vimp_summary)
}

list
# After loop
cindex_all <- sapply(list, function(x) x$cindex)
cindex_all
mean(cindex_all)
sd(cindex_all)/sqrt(4)
list
# Aggregate variable importance
all_vimp <- do.call(cbind, lapply(list, function(x) x$vimp$Mean))
rowMeans(all_vimp, na.rm = TRUE)
all_vimp
vimp_summary <- data.frame(
  Variable = list[[1]]$vimp$Variable,
  Mean     = rowMeans(all_vimp),
  SD       = apply(all_vimp, 1, sd)
)

vimp_summary
```


### model 6
```{r}
seed <- sample(seq(from = 1, to = 10000), size = 30, replace = FALSE)
saveRDS(seeds, "seed_7.txt")
list = NULL
for (s in seeds) {
  set.seed(s)
  folds <- createFolds(clean$Recurrence, k = 4, returnTrain = TRUE)
  cindex_vals <- numeric(length(folds))
  vimp_list <- vector("list", length(folds))
  
  for (j in seq_along(folds)) {
    train_data <- clean[folds[[j]], ]
    test_data  <- clean[-folds[[j]], ]
    
    mod <- rfsrc(
     Surv(Time_to_event, Recurrence) ~ Mole_type_tri + Histology + T_stage_binary + Surgery_type + N_stage + Antiher2 + Endocrine + Oophorectomy + Pathologic_response + Patho_N_binary + Near_response_strict,
      data = train_data,
      ntree = 1000,
      na.action = "na.impute"
    )
    
    vimp_vals <- vimp(mod)$importance
    vimp_list[[j]] <- vimp_vals
    
    pred <- predict(mod, newdata = test_data, na.action = "na.impute")
    risk_scores <- pred$predicted
    surv_obj <- Surv(test_data$Time_to_event, test_data$Recurrence)
    cindex_vals[j] <- survConcordance(surv_obj ~ risk_scores)$concordance
  }
  
  # Store average c-index and VIMP summary
  mean_cindex <- mean(cindex_vals)
  print(mean_cindex)
  
  vimp_df <- do.call(cbind, vimp_list)
  vimp_summary <- data.frame(
    Variable = rownames(vimp_df),
    Mean = rowMeans(vimp_df, na.rm = TRUE),
    SD = apply(vimp_df, 1, sd, na.rm = TRUE)
  )
  
  vimp_summary <- vimp_summary %>% arrange(Mean)
  list[[as.character(s)]] <- list(cindex = mean_cindex, vimp = vimp_summary)
}

list
# After loop
cindex_all <- sapply(list, function(x) x$cindex)
cindex_all
mean(cindex_all)
sd(cindex_all)/sqrt(4)
list
# Aggregate variable importance
all_vimp <- do.call(cbind, lapply(list, function(x) x$vimp$Mean))
rowMeans(all_vimp, na.rm = TRUE)
all_vimp
vimp_summary <- data.frame(
  Variable = list[[1]]$vimp$Variable,
  Mean     = rowMeans(all_vimp),
  SD       = apply(all_vimp, 1, sd)
)

vimp_summary
```


### model 8
```{r}
seed <- sample(seq(from = 1, to = 10000), size = 30, replace = FALSE)
saveRDS(seeds, "seed_9.txt")
list = NULL
for (s in seeds) {
  set.seed(s)
  folds <- createFolds(clean$Recurrence, k = 4, returnTrain = TRUE)
  cindex_vals <- numeric(length(folds))
  vimp_list <- vector("list", length(folds))
  
  for (j in seq_along(folds)) {
    train_data <- clean[folds[[j]], ]
    test_data  <- clean[-folds[[j]], ]
    
    mod <- rfsrc(
     Surv(Time_to_event, Recurrence) ~  Race_binary + Mole_type_tri + Nottingham_score_binary + Surgery + Surgery_type + 
      Chemo_adjuvent + Antiher2_treatment + Endocrine + Oophorectomy + Ovarian_status + 
      Neoadjuvant + Pathologic_response + Patho_N_binary + 
      Patho_T_binary + Patho_M_Binary + Near_response_strict_binary,
      data = train_data,
      ntree = 1000,
      na.action = "na.impute"
    )
    
    vimp_vals <- vimp(mod)$importance
    vimp_list[[j]] <- vimp_vals
    
    pred <- predict(mod, newdata = test_data, na.action = "na.impute")
    risk_scores <- pred$predicted
    surv_obj <- Surv(test_data$Time_to_event, test_data$Recurrence)
    cindex_vals[j] <- survConcordance(surv_obj ~ risk_scores)$concordance
  }
  
  # Store average c-index and VIMP summary
  mean_cindex <- mean(cindex_vals)
  print(mean_cindex)
  
  vimp_df <- do.call(cbind, vimp_list)
  vimp_summary <- data.frame(
    Variable = rownames(vimp_df),
    Mean = rowMeans(vimp_df, na.rm = TRUE),
    SD = apply(vimp_df, 1, sd, na.rm = TRUE)
  )
  
  vimp_summary <- vimp_summary %>% arrange(Mean)
  list[[as.character(s)]] <- list(cindex = mean_cindex, vimp = vimp_summary)
}

list
# After loop
cindex_all_8 <- sapply(list, function(x) x$cindex)
cindex_all
mean(cindex_all_8)
sd(cindex_all_8)/sqrt(4)

# Aggregate variable importance
all_vimp_8 <- do.call(cbind, lapply(list, function(x) x$vimp$Mean))
rowMeans(all_vimp_8, na.rm = TRUE)
all_vimp_8
vimp_summary_8 <- data.frame(
  Variable = list[[1]]$vimp$Variable,
  Mean     = rowMeans(all_vimp),
  SD       = apply(all_vimp, 1, sd)
)

vimp_summary_8
```

model 9 (original binary model)
```{r}
seed <- sample(seq(from = 1, to = 10000), size = 30, replace = FALSE)
saveRDS(seeds, "seed_9.txt")
list = NULL
for (s in seeds) {
  set.seed(s)
  folds <- createFolds(clean$Recurrence, k = 4, returnTrain = TRUE)
  cindex_vals <- numeric(length(folds))
  vimp_list <- vector("list", length(folds))
  
  for (j in seq_along(folds)) {
    train_data <- clean[folds[[j]], ]
    test_data  <- clean[-folds[[j]], ]
    
    mod <- rfsrc(
     Surv(Time_to_event, Recurrence) ~  Race + Mole_type + Nottingham_score + Surgery + Surgery_type + 
      Chemo + Antiher2 + Endocrine + Oophorectomy + Ovarian_status + 
      Neoadjuvant + Pathologic_response + Patho_N_stage + 
      Patho_T_stage + Patho_M_stage,
      data = train_data,
      ntree = 1000,
      na.action = "na.impute"
    )
    
    vimp_vals <- vimp(mod)$importance
    vimp_list[[j]] <- vimp_vals
    
    pred <- predict(mod, newdata = test_data, na.action = "na.impute")
    risk_scores <- pred$predicted
    surv_obj <- Surv(test_data$Time_to_event, test_data$Recurrence)
    cindex_vals[j] <- survConcordance(surv_obj ~ risk_scores)$concordance
  }
  
  # Store average c-index and VIMP summary
  mean_cindex <- mean(cindex_vals)
  print(mean_cindex)
  
  vimp_df <- do.call(cbind, vimp_list)
  vimp_summary <- data.frame(
    Variable = rownames(vimp_df),
    Mean = rowMeans(vimp_df, na.rm = TRUE),
    SD = apply(vimp_df, 1, sd, na.rm = TRUE)
  )
  
  vimp_summary <- vimp_summary %>% arrange(Mean)
  list[[as.character(s)]] <- list(cindex = mean_cindex, vimp = vimp_summary)
}

list
# After loop
cindex_all_9 <- sapply(list, function(x) x$cindex)
cindex_all_9
mean(cindex_all_9)
sd(cindex_all_9)/sqrt(4)

# Aggregate variable importance
all_vimp <- do.call(cbind, lapply(list, function(x) x$vimp$Mean))
rowMeans(all_vimp, na.rm = TRUE)
all_vimp
vimp_summary_9 <- data.frame(
  Variable = list[[1]]$vimp$Variable,
  Mean     = rowMeans(all_vimp),
  SD       = apply(all_vimp, 1, sd)
)

vimp_summary_9
```

## combinations of orginial and dichotomous variables

### model 11 (combo model binary model)
```{r}
seed <- sample(seq(from = 1, to = 10000), size = 30, replace = FALSE)
saveRDS(seeds, "seed_11.txt")
list = NULL
for (s in seeds) {
  set.seed(s)
  folds <- createFolds(clean$Recurrence, k = 4, returnTrain = TRUE)
  cindex_vals <- numeric(length(folds))
  vimp_list <- vector("list", length(folds))
  
  for (j in seq_along(folds)) {
    train_data <- clean[folds[[j]], ]
    test_data  <- clean[-folds[[j]], ]
    
    mod <- rfsrc(
     Surv(Time_to_event, Recurrence) ~  Mole_type_tri + Nottingham_score_binary + 
      Chemo + Antiher2 + Endocrine + Oophorectomy  + 
      Neoadjuvant + Pathologic_response + Patho_N_stage + 
      Patho_T_stage + Patho_M_Binary + Near_response_strict_binary,
      data = train_data,
      ntree = 1000,
      na.action = "na.impute"
    )
    
    vimp_vals <- vimp(mod)$importance
    vimp_list[[j]] <- vimp_vals
    
    pred <- predict(mod, newdata = test_data, na.action = "na.impute")
    risk_scores <- pred$predicted
    surv_obj <- Surv(test_data$Time_to_event, test_data$Recurrence)
    cindex_vals[j] <- survConcordance(surv_obj ~ risk_scores)$concordance
  }
  
  # Store average c-index and VIMP summary
  mean_cindex <- mean(cindex_vals)
  print(mean_cindex)
  
  vimp_df <- do.call(cbind, vimp_list)
  vimp_summary <- data.frame(
    Variable = rownames(vimp_df),
    Mean = rowMeans(vimp_df, na.rm = TRUE),
    SD = apply(vimp_df, 1, sd, na.rm = TRUE)
  )
  
  vimp_summary <- vimp_summary %>% arrange(Mean)
  list[[as.character(s)]] <- list(cindex = mean_cindex, vimp = vimp_summary)
}

list
# After loop
cindex_all_11 <- sapply(list, function(x) x$cindex)
cindex_all_11
mean(cindex_all_11)
sd(cindex_all_11)/sqrt(4)

# Aggregate variable importance
all_vimp_11 <- do.call(cbind, lapply(list, function(x) x$vimp$Mean))
rowMeans(all_vimp_11, na.rm = TRUE)
all_vimp_11
vimp_summary_11 <- data.frame(
  Variable = list[[1]]$vimp$Variable,
  Mean     = rowMeans(all_vimp_11),
  SD       = apply(all_vimp_11, 1, sd)
)

vimp_summary_11
```

### model 12 (combo model binary model)
```{r}
seed <- sample(seq(from = 1, to = 10000), size = 30, replace = FALSE)
saveRDS(seeds, "seed_11.txt")
list = NULL
for (s in seeds) {
  set.seed(s)
  folds <- createFolds(clean$Recurrence, k = 4, returnTrain = TRUE)
  cindex_vals <- numeric(length(folds))
  vimp_list <- vector("list", length(folds))
  
  for (j in seq_along(folds)) {
    train_data <- clean[folds[[j]], ]
    test_data  <- clean[-folds[[j]], ]
    
    mod <- rfsrc(
     Surv(Time_to_event, Recurrence) ~  Mole_type + Nottingham_score + 
      Chemo + Antiher2 + Endocrine + Pathologic_response + Patho_N_stage + 
      Patho_T_stage + Patho_M_Binary + Near_response_strict_binary,
      data = train_data,
      ntree = 1000,
      na.action = "na.impute"
    )
    
    vimp_vals <- vimp(mod)$importance
    vimp_list[[j]] <- vimp_vals
    
    pred <- predict(mod, newdata = test_data, na.action = "na.impute")
    risk_scores <- pred$predicted
    surv_obj <- Surv(test_data$Time_to_event, test_data$Recurrence)
    cindex_vals[j] <- survConcordance(surv_obj ~ risk_scores)$concordance
  }
  
  # Store average c-index and VIMP summary
  mean_cindex <- mean(cindex_vals)
  print(mean_cindex)
  
  vimp_df <- do.call(cbind, vimp_list)
  vimp_summary <- data.frame(
    Variable = rownames(vimp_df),
    Mean = rowMeans(vimp_df, na.rm = TRUE),
    SD = apply(vimp_df, 1, sd, na.rm = TRUE)
  )
  
  vimp_summary <- vimp_summary %>% arrange(Mean)
  list[[as.character(s)]] <- list(cindex = mean_cindex, vimp = vimp_summary)
}

list
# After loop
cindex_all_12 <- sapply(list, function(x) x$cindex)
cindex_all_12
mean(cindex_all_12)
sd(cindex_all_12)/sqrt(4)

# Aggregate variable importance
all_vimp_12 <- do.call(cbind, lapply(list, function(x) x$vimp$Mean))
rowMeans(all_vimp_12, na.rm = TRUE)
all_vimp_11
vimp_summary_12 <- data.frame(
  Variable = list[[1]]$vimp$Variable,
  Mean     = rowMeans(all_vimp_12),
  SD       = apply(all_vimp_12, 1, sd)
)

vimp_summary_12
```


### model 13 (combo model binary model)
```{r}
seed <- sample(seq(from = 1, to = 10000), size = 30, replace = FALSE)
saveRDS(seeds, "seed_13.txt")
list = NULL
for (s in seeds) {
  set.seed(s)
  folds <- createFolds(clean$Recurrence, k = 4, returnTrain = TRUE)
  cindex_vals <- numeric(length(folds))
  vimp_list <- vector("list", length(folds))
  
  for (j in seq_along(folds)) {
    train_data <- clean[folds[[j]], ]
    test_data  <- clean[-folds[[j]], ]
    
    mod <- rfsrc(
     Surv(Time_to_event, Recurrence) ~  Mole_type + Nottingham_score + 
      Chemo + Antiher2 + Endocrine + Pathologic_response + Patho_N_stage + 
      Patho_T_stage + Patho_M_Binary + Near_response_strict_binary,
      data = train_data,
      ntree = 1000,
      na.action = "na.impute"
    )
    
    vimp_vals <- vimp(mod)$importance
    vimp_list[[j]] <- vimp_vals
    
    pred <- predict(mod, newdata = test_data, na.action = "na.impute")
    risk_scores <- pred$predicted
    surv_obj <- Surv(test_data$Time_to_event, test_data$Recurrence)
    cindex_vals[j] <- survConcordance(surv_obj ~ risk_scores)$concordance
  }
  
  # Store average c-index and VIMP summary
  mean_cindex <- mean(cindex_vals)
  print(mean_cindex)
  
  vimp_df <- do.call(cbind, vimp_list)
  vimp_summary <- data.frame(
    Variable = rownames(vimp_df),
    Mean = rowMeans(vimp_df, na.rm = TRUE),
    SD = apply(vimp_df, 1, sd, na.rm = TRUE)
  )
  
  vimp_summary <- vimp_summary %>% arrange(Mean)
  list[[as.character(s)]] <- list(cindex = mean_cindex, vimp = vimp_summary)
}

list
# After loop
cindex_all_13 <- sapply(list, function(x) x$cindex)
cindex_all_13
mean(cindex_all_13)
sd(cindex_all_13)/sqrt(4)

# Aggregate variable importance
all_vimp_13 <- do.call(cbind, lapply(list, function(x) x$vimp$Mean))
rowMeans(all_vimp_13, na.rm = TRUE)
all_vimp_13
vimp_summary_13 <- data.frame(
  Variable = list[[1]]$vimp$Variable,
  Mean     = rowMeans(all_vimp_13),
  SD       = apply(all_vimp_13, 1, sd)
)

vimp_summary_13
```


### model 14 second best (0.7527296)
```{r}
seed <- sample(seq(from = 1, to = 10000), size = 30, replace = FALSE)
saveRDS(seeds, "seed_14.txt")
list = NULL
for (s in seeds) {
  set.seed(s)
  folds <- createFolds(clean$Recurrence, k = 4, returnTrain = TRUE)
  cindex_vals <- numeric(length(folds))
  vimp_list <- vector("list", length(folds))
  
  for (j in seq_along(folds)) {
    train_data <- clean[folds[[j]], ]
    test_data  <- clean[-folds[[j]], ]
    
    mod <- rfsrc(
     Surv(Time_to_event, Recurrence) ~  Mole_type_tri + Histology + T_stage_binary +
       Surgery_type + N_stage + Antiher2 + Endocrine + Oophorectomy + Pathologic_response + Patho_N_binary + Near_response_strict,
      data = train_data,
      ntree = 1000,
      na.action = "na.impute"
    )
    
    vimp_vals <- vimp(mod)$importance
    vimp_list[[j]] <- vimp_vals
    
    pred <- predict(mod, newdata = test_data, na.action = "na.impute")
    risk_scores <- pred$predicted
    surv_obj <- Surv(test_data$Time_to_event, test_data$Recurrence)
    cindex_vals[j] <- survConcordance(surv_obj ~ risk_scores)$concordance
  }
  
  # Store average c-index and VIMP summary
  mean_cindex <- mean(cindex_vals)
  print(mean_cindex)
  
  vimp_df <- do.call(cbind, vimp_list)
  vimp_summary <- data.frame(
    Variable = rownames(vimp_df),
    Mean = rowMeans(vimp_df, na.rm = TRUE),
    SD = apply(vimp_df, 1, sd, na.rm = TRUE)
  )
  
  vimp_summary <- vimp_summary %>% arrange(Mean)
  list[[as.character(s)]] <- list(cindex = mean_cindex, vimp = vimp_summary)
}

list
# After loop
cindex_all_14 <- sapply(list, function(x) x$cindex)
cindex_all_14
mean(cindex_all_14)
sd(cindex_all_14)/sqrt(4)

# Aggregate variable importance
all_vimp_14 <- do.call(cbind, lapply(list, function(x) x$vimp$Mean))
rowMeans(all_vimp_14, na.rm = TRUE)
all_vimp_14
vimp_summary_14 <- data.frame(
  Variable = list[[1]]$vimp$Variable,
  Mean     = rowMeans(all_vimp_14),
  SD       = apply(all_vimp_14, 1, sd)
)

vimp_summary_14
```

### model 15 (best 0.7554698)

```{r}
seed <- sample(seq(from = 1, to = 10000), size = 30, replace = FALSE)
saveRDS(seeds, "seed_15.txt")
list = NULL
for (s in seeds) {
  set.seed(s)
  folds <- createFolds(clean$Recurrence, k = 4, returnTrain = TRUE)
  cindex_vals <- numeric(length(folds))
  vimp_list <- vector("list", length(folds))
  
  for (j in seq_along(folds)) {
    train_data <- clean[folds[[j]], ]
    test_data  <- clean[-folds[[j]], ]
    
    mod <- rfsrc(
     Surv(Time_to_event, Recurrence) ~  Mole_type_tri + Histology  +
       Surgery_type + N_stage + Antiher2 + Endocrine + Oophorectomy + Pathologic_response + Patho_N_binary + Near_response_strict,
      data = train_data,
      ntree = 1000,
      na.action = "na.impute"
    )
    
    vimp_vals <- vimp(mod)$importance
    vimp_list[[j]] <- vimp_vals
    
    pred <- predict(mod, newdata = test_data, na.action = "na.impute")
    risk_scores <- pred$predicted
    surv_obj <- Surv(test_data$Time_to_event, test_data$Recurrence)
    cindex_vals[j] <- survConcordance(surv_obj ~ risk_scores)$concordance
  }
  
  # Store average c-index and VIMP summary
  mean_cindex <- mean(cindex_vals)
  print(mean_cindex)
  
  vimp_df <- do.call(cbind, vimp_list)
  vimp_summary <- data.frame(
    Variable = rownames(vimp_df),
    Mean = rowMeans(vimp_df, na.rm = TRUE),
    SD = apply(vimp_df, 1, sd, na.rm = TRUE)
  )
  
  vimp_summary <- vimp_summary %>% arrange(Mean)
  list[[as.character(s)]] <- list(cindex = mean_cindex, vimp = vimp_summary)
}

list
# After loop
cindex_all_15 <- sapply(list, function(x) x$cindex)
cindex_all_15
mean(cindex_all_15)
sd(cindex_all_15)/sqrt(4)

# Aggregate variable importance
all_vimp_15 <- do.call(cbind, lapply(list, function(x) x$vimp$Mean))
rowMeans(all_vimp_15, na.rm = TRUE)
all_vimp_15
vimp_summary_15 <- data.frame(
  Variable = list[[1]]$vimp$Variable,
  Mean     = rowMeans(all_vimp_15),
  SD       = apply(all_vimp_15, 1, sd)
)

vimp_summary_15
```

### 16 (removed surgery type not as good .7488518)

```{r}
seed <- sample(seq(from = 1, to = 10000), size = 30, replace = FALSE)
saveRDS(seeds, "seed_16.txt")
list = NULL
for (s in seeds) {
  set.seed(s)
  folds <- createFolds(clean$Recurrence, k = 4, returnTrain = TRUE)
  cindex_vals <- numeric(length(folds))
  vimp_list <- vector("list", length(folds))
  
  for (j in seq_along(folds)) {
    train_data <- clean[folds[[j]], ]
    test_data  <- clean[-folds[[j]], ]
    
    mod <- rfsrc(
     Surv(Time_to_event, Recurrence) ~  Mole_type_tri + Histology + N_stage + Antiher2 + Endocrine + Oophorectomy + Pathologic_response + Patho_N_binary + Near_response_strict,
      data = train_data,
      ntree = 1000,
      na.action = "na.impute"
    )
    
    vimp_vals <- vimp(mod)$importance
    vimp_list[[j]] <- vimp_vals
    
    pred <- predict(mod, newdata = test_data, na.action = "na.impute")
    risk_scores <- pred$predicted
    surv_obj <- Surv(test_data$Time_to_event, test_data$Recurrence)
    cindex_vals[j] <- survConcordance(surv_obj ~ risk_scores)$concordance
  }
  
  # Store average c-index and VIMP summary
  mean_cindex <- mean(cindex_vals)
  print(mean_cindex)
  
  vimp_df <- do.call(cbind, vimp_list)
  vimp_summary <- data.frame(
    Variable = rownames(vimp_df),
    Mean = rowMeans(vimp_df, na.rm = TRUE),
    SD = apply(vimp_df, 1, sd, na.rm = TRUE)
  )
  
  vimp_summary <- vimp_summary %>% arrange(Mean)
  list[[as.character(s)]] <- list(cindex = mean_cindex, vimp = vimp_summary)
}

list
# After loop
cindex_all_16 <- sapply(list, function(x) x$cindex)
cindex_all_16
mean(cindex_all_16)
sd(cindex_all_16)/sqrt(4)

# Aggregate variable importance
all_vimp_16 <- do.call(cbind, lapply(list, function(x) x$vimp$Mean))
rowMeans(all_vimp_16, na.rm = TRUE)
all_vimp_16
vimp_summary_16 <- data.frame(
  Variable = list[[1]]$vimp$Variable,
  Mean     = rowMeans(all_vimp_16),
  SD       = apply(all_vimp_16, 1, sd)
)

vimp_summary_16
```

### model 17 (removed histology new best 0.7713017)

```{r}
seeds <- sample(seq(from = 1, to = 10000), size = 30, replace = FALSE)
saveRDS(seeds, "seed_17.txt")
list = NULL
for (s in seeds) {
  set.seed(s)
  folds <- createFolds(clean$Recurrence, k = 4, returnTrain = TRUE)
  cindex_vals <- numeric(length(folds))
  vimp_list <- vector("list", length(folds))
  
  for (j in seq_along(folds)) {
    train_data <- clean[folds[[j]], ]
    test_data  <- clean[-folds[[j]], ]
    
    mod <- rfsrc(
     Surv(Time_to_event, Recurrence) ~  Mole_type_tri + Surgery_type + N_stage + Antiher2_treatment + Endocrine + Oophorectomy + Pathologic_response + Patho_N_binary + Near_response_loose,
      data = train_data,
      ntree = 1500,
      na.action = "na.imput",
     nodesize = 1,
     mtry = 2
     
    )
    
    vimp_vals <- vimp(mod)$importance
    vimp_list[[j]] <- vimp_vals
    
    pred <- predict.rfsrc(mod, newdata = test_data, na.action = "na.imput")
    risk_scores <- pred$predicted
    surv_obj <- Surv(test_data$Time_to_event, test_data$Recurrence)
    cindex_vals[j] <- survConcordance(surv_obj ~ risk_scores)$concordance
  }
  
  # Store average c-index and VIMP summary
  mean_cindex <- mean(cindex_vals)
  print(mean_cindex)
  
  vimp_df <- do.call(cbind, vimp_list)
  vimp_summary <- data.frame(
    Variable = rownames(vimp_df),
    Mean = rowMeans(vimp_df, na.rm = TRUE),
    SD = apply(vimp_df, 1, sd, na.rm = TRUE)
  )
  
  vimp_summary <- vimp_summary %>% arrange(Mean)
  list[[as.character(s)]] <- list(cindex = mean_cindex, vimp = vimp_summary)
}

list
# After loop
cindex_all_17 <- sapply(list, function(x) x$cindex)
cindex_all_17
mean <- mean(cindex_all_17)
mean
se <- sd(cindex_all_17)/sqrt(100)
nine5ci <- mean + se * c(-1.96, 1.96)
nine5ci
# Aggregate variable importance
all_vimp_17 <- do.call(cbind, lapply(list, function(x) x$vimp$Mean))
rowMeans(all_vimp_17, na.rm = TRUE)
all_vimp_17
vimp_summary_17 <- data.frame(
  Variable = list[[1]]$vimp$Variable,
  Mean     = rowMeans(all_vimp_17),
  SD       = apply(all_vimp_17, 1, sd)
)

vimp_summary_17
```

# best model and comparisons

### 17 changing my predict (removed histology new best 0.782874)-flagship model
```{r}
#seeds <- sample(seq(from = 1, to = 10000), size = 100, replace = FALSE)
df <- clean %>% select(mole_type_three, Surgery_type, N_stage, Antiher2, Endocrine, Oophorectomy, Pathologic_response, Patho_N_binary, Near_response_strict)
vis_miss(df, sort_miss = TRUE)
sum(complete.cases(df))
list = NULL
seeds <- readRDS("seed_17.rmd")
for (s in seeds) {
  set.seed(s)
  folds <- createFolds(clean$Recurrence, k = 4, returnTrain = TRUE) 
  cindex_vals <- numeric(length(folds))
  vimp_list <- vector("list", length(folds))

  for (j in seq_along(folds)) {
    test_data <- clean[-folds[[j]], ]
    train_data  <- clean[folds[[j]], ]
    
    mod <- rfsrc(
     Surv(Time_to_event, Recurrence) ~  mole_type_three + Surgery_type + N_stage + Antiher2 + Endocrine + Oophorectomy + Near_response_strict + T_stage,
      data = train_data,
      ntree = 4000,
      na.action = "na.imput",
     nodesize = 1,
     mtry = 2,
    )
  
    vimp_vals <- vimp(mod)$importance
    vimp_list[[j]] <- vimp_vals
    
    pred <- predict.rfsrc(mod, newdata = test_data, na.action = "na.imput")
    risk_scores <- pred$predicted
    surv_obj <- Surv(test_data$Time_to_event, test_data$Recurrence)
    cindex_vals[j] <- survConcordance(surv_obj ~ risk_scores)$concordance
  }
  
  # Store average c-index and VIMP summary
  mean_cindex <- mean(cindex_vals)
  print(mean_cindex)
  
  vimp_df <- do.call(cbind, vimp_list)
  vimp_summary <- data.frame(
    Variable = rownames(vimp_df),
    Mean = rowMeans(vimp_df, na.rm = TRUE),
    SD = apply(vimp_df, 1, sd, na.rm = TRUE)
  )
  
  vimp_summary <- vimp_summary %>% arrange(Mean)
  list[[as.character(s)]] <- list(cindex = mean_cindex, vimp = vimp_summary)
}

list
# After loop
cindex_all_17_v2_fix <- sapply(list, function(x) x$cindex)
cindex_all_17_v2_fix
mean <- mean(cindex_all_17_v2_fix)
mean
se <- sd(cindex_all_17_v2_fix)/sqrt(100)
nine5ci <- mean + se * c(-1.96, 1.96)
nine5ci
# Aggregate variable importance
all_vimp_17_v2_fix <- do.call(cbind, lapply(list, function(x) x$vimp$Mean))
rowMeans(all_vimp_17_v2_fix, na.rm = TRUE)
all_vimp_17_v2_fix
vimp_summary_17_v2_fix <- data.frame(
  Variable = list[[1]]$vimp$Variable,
  Mean     = rowMeans(all_vimp_17_v2_fix),
  SD       = apply(all_vimp_17_v2_fix, 1, sd)
)
#[1] 0.7754595, [1] 0.7712222 0.7796968

vimp_summary_17_v2_fix
#0.7790911-0.7833892

```

### original model
```{r}
#seeds <- sample(seq(from = 1, to = 10000), size = 100, replace = FALSE)
seeds <- readRDS("seed_17.rmd")
list = NULL
seeds
for (s in seeds) {
  set.seed(s)
  folds <- createFolds(clean$Recurrence, k = 4, returnTrain = TRUE)
  cindex_vals <- numeric(length(folds))
  vimp_list <- vector("list", length(folds))
  
  for (j in seq_along(folds)) {
    train_data <- clean[folds[[j]], ]
    test_data  <- clean[-folds[[j]], ]
   train_data$dummy <- 1
   test_data$dummy <- 1
    mod <- rfsrc(
     Surv(Time_to_event, Recurrence) ~  Race + Mole_type + T_stage + N_stage + Surgery_type + 
      Chemo + Antiher2 + Endocrine + Oophorectomy + Ovarian_status + Pathologic_response + Patho_N_stage +
      Patho_T_stage + Patho_M_stage + Near_response_strict,
      data = train_data,
      ntree = 4000,
      na.action = "na.imput"
    )
  
    vimp_vals <- vimp(mod)$importance
    vimp_list[[j]] <- vimp_vals
    
    pred <- predict.rfsrc(mod, newdata = test_data, na.action = "na.imput")
    risk_scores <- pred$predicted
    surv_obj <- Surv(test_data$Time_to_event, test_data$Recurrence)
    cindex_vals[j] <- survConcordance(surv_obj ~ risk_scores)$concordance
  }
  
  # Store average c-index and VIMP summary
  mean_cindex <- mean(cindex_vals)
  print(mean_cindex)
  
  vimp_df <- do.call(cbind, vimp_list)
  vimp_summary <- data.frame(
  Variable = rownames(vimp_df),
  Mean = rowMeans(vimp_df, na.rm = TRUE),
  SD = apply(vimp_df, 1, sd, na.rm = TRUE)
  )
  
  list[[as.character(s)]] <- list(cindex = mean_cindex)
}

list
# After loop
cindex_all_og <- sapply(list, function(x) x$cindex)
cindex_all_og
mean <- mean(cindex_all_og)
mean
se <- sd(cindex_all_og)/sqrt(100)
nine5ci <- mean + se * c(-1.96, 1.96)
nine5ci
# Aggregate variable importance
all_vimp_og <- do.call(cbind, lapply(list, function(x) x$vimp$Mean))
rowMeans(all_vimp_og, na.rm = TRUE)
all_vimp_og
vimp_summary_og <- data.frame(
  Variable = list[[1]]$vimp$Variable,
  Mean     = rowMeans(all_vimp_og),
  SD       = apply(all_vimp_og, 1, sd)
)

vimp_summary_og
#[1] 0.7367926, 0.7324617 0.7411234 
#0.7249231-0.7204966

cindex.comp(0.7367926, 0.7754595)


get.tree(mod, 1)

library(compareC)

compareC(time = clean$Time_to_event,
         status = clean$Recurrence,
         score1 = 0.7367926,
         score2 = 0.7754595)
```

## model 17 (graphing prediction for a specific subgroup) 
```{r}
library(scales)

set.seed(8)
 mod <- rfsrc(
     Surv(Time_to_event, Recurrence) ~  mole_type_three + Surgery_type + N_stage + Antiher2 + Endocrine + Oophorectomy + Near_response_strict,
      data = train_data,
      ntree = 4000,
      na.action = "na.imput",
     nodesize = 1,
     mtry = 2,
     k = 4
    )
plot <- plot.survival(mod, show.plots = TRUE)
smp.o <- subsample(mod, B=25,
subratio=.5)
plot.subsample(smp.o, sorted = TRUE)

smp.o
```

# graphs and plots

## node plot 
```{r}
library(party)
surv_tree <- ctree(Surv(Time_to_event, Recurrence) ~  mole_type_three + Surgery_type + N_stage + Antiher2 + Endocrine + Oophorectomy + Near_response_strict, 
                   data = clean)
png("ctree_with_KM.png", width = 1600, height = 1000, res = 200)
plot(surv_tree)  # Keeps KM curves
dev.off()

cleaned_data %>%
  filter(Surgery == "1")
```

## km plot graph
```{r}
km_plot <- survfit(Surv(Time_to_event/365, Recurrence) ~ mole_type_three, data = clean)
library(survival)
ttle <- paste0("Time to Reccurence by ","Receptor Status")
library(survminer)
library(ggplot2)
library(ggfortify)
# Build the plot
the_plot <- autoplot(km_plot, conf.int = FALSE) +  # remove confidence intervals
  labs(
    col = ttle,
    fill = ttle,
    title = ttle,
    x = "Time (years)",
    y = "Recurrence-Free Probability"
  ) +
  theme_light(base_size = 16) +  # overall font size
  theme(
    legend.position = "bottom",
    plot.title = element_text(size = 26, face = "bold"),
    legend.title = element_text(size = 14, hjust = 0),
    legend.text = element_text(size = 14, hjust = 0),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14)
  ) +
  scale_color_brewer(palette = "Dark2") +    # colorblind-friendly
  scale_fill_brewer(palette = "Dark2")       # colorblind-friendly
the_plot
ggsave("survival_plot_square.png", plot = the_plot, width = 9, height = 5)

```


## graphs of neoadjuvent models 
```{r}
box_data <- data.frame(
  group = c("Receptor Status", "Surgery Type", "N Stage", "Antiher-2 Therapy", 
            "Endocrine Therapy", "Oophorectomy", "Near Complete Response"),
  ymin = c(-0.004 , -0.030,  0.006, 0.025,  -0.001,  0.012, 0.092),      # min or lower error bound
  middle = c(0.067,  0.039, 0.108,  0.096, 0.075, 0.090,  0.210),        # main value
  ymax = c( 0.138, 0.108, 0.210, 0.167, 0.151, 0.168, 0.328)             # max or upper error bound
) %>% 
  mutate(significance = ifelse(ymin >= 0, "Significant", "Not Significant")) %>%
  arrange(desc(middle))



ggplot(data = box_data, aes(x = reorder(group, middle), y = middle, fill = significance)) + 
  geom_col(width = 0.6) +  # bar height = 'middle'
  geom_errorbar(aes(ymin = ymin, ymax = ymax), width = 0.2, linewidth = 0.6) +  # error bars
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +  # zero reference line
  scale_fill_manual(values = c("Significant" = "#FFCB05", "Not Significant" = "#648FFF"),
                    name = "Variable Significance") +
  scale_x_discrete(labels = wrap_format(10)) +
  labs(
    title = "The Importance of Variables Used",
    x = "Variables",
    y = "Variable Importance (VIMP)"
  ) +
  coord_cartesian(ylim = c(-0.1, .4)) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
    axis.title.x = element_text(size = 18, face = "bold"),
    axis.title.y = element_text(size = 15, face ="bold"),
    axis.text.y = element_text(size = 14, color = "black"),
    axis.text.x = element_text(size = 14, color = "black"),
    legend.text =  element_text(size = 14, color = "black"),
    legend.title =  element_text(size = 14, color = "black", face = "bold")
  )
ggsave("Figures/vimp.jpg", units = "in", width = 9, height = 5, dpi = 300)
```


## risk curve 1
```{r}
risk_curve = NULL
for(i in 0:3) {
  val <- as.character(i)
  
  person <- data.frame(
  mole_type_three        = "Triple Negative",                   # "HER+", "Luminal-like", "Triple Negative"
  Surgery_type         = "0",                                 # levels are "0","1"; so keep it as a string
  N_stage              = i,                                 # levels are "0","1","2","3"
  Antiher2             = "None",                              # Adjuvant Only Both Neoadjuvant Only None
  Endocrine            = "None",                              # Adjuvant Only, Both, Neoadjuvant Only, None
  Oophorectomy         = "No",                                # "No" / "Yes"
  Near_response_strict = "Not complete",                      # "Not complete" "Complete" "Near-complete"
  stringsAsFactors = FALSE
)
  
for (v in names(person)) {
  person[[v]] <- factor(person[[v]],
                        levels = levels(mod$xvar[[v]]))
}
  pred <- predict(mod, newdata = person)
curve_df <- data.frame(time = pred$time.interest,
                       surv = pred$survival[1, ])
curve_df$T_stage <- i

  name <- paste0("person_", i)
  risk_curve <- rbind2(risk_curve, curve_df)

  assign(x = "risk_curve", value = risk_curve, envir = .GlobalEnv)
}
risk_curve
ggplot(risk_curve, aes(time / 365, surv, color = as.character(T_stage), group = as.character(T_stage))) + 
  scale_color_manual(values = c("0" = "#D81B60", "1" = "#1E88E5", "2" = "#FFC107", "3" = "#004D40"), 
                    labels = c("N0", "N1", "N2", "N3"), 
                    name = "N stage") +
  geom_step(linewidth = 1) +
  labs(x = "Time (years)", y = "Recurrence-Free Probability",
       title = "Predicted Recurrence-Free Curve for Triple Negative Cancer") + 
  scale_y_continuous(limits = c(.2, 1), expand = c(0, 0)) + 
  theme_minimal(base_size = 15) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
    axis.title.x = element_text(size = 18, face = "bold"),
    axis.title.y = element_text(size = 15, face ="bold"),
    axis.text.y = element_text(size = 14, color = "black"),
    axis.text.x = element_text(size = 14, color = "black"),
    legend.text =  element_text(size = 14, color = "black"),
    legend.title =  element_text(size = 14, color = "black", face = "bold"))

ggsave("Figures/triple_neg.jpg", units = "in", width = 7, height = 7, dpi = 300)
```

## risk curve 2

```{r}
str(person)
risk_curve = NULL
for(i in 0:3) {
  val <- as.character(i)
  
  person <- data.frame(
  mole_type_three        = "PR+ or ER+",                   # "HER+", "Luminal-like", "Triple Negative"
  Surgery_type         = "0",                                 # levels are "0","1"; so keep it as a string
  N_stage              = i,                                 # levels are "0","1","2","3"
  Antiher2             = "None",                              # Adjuvant Only Both Neoadjuvant Only None
  Endocrine            = "Neoadjuvant Only",                  # Adjuvant Only, Both, Neoadjuvant Only, None
  Oophorectomy         = "No",                                # "No" / "Yes"
  Near_response_strict = "Not complete",                      # "Not complete" "Complete" "Near-complete"
  stringsAsFactors = FALSE
)
  
for (v in names(person)) {
  person[[v]] <- factor(person[[v]],
                        levels = levels(mod$xvar[[v]]))
}
  pred <- predict(mod, newdata = person)
curve_df <- data.frame(time = pred$time.interest,
                       surv = pred$survival[1, ])
curve_df$T_stage <- i

  name <- paste0("person_", i)
  risk_curve <- rbind2(risk_curve, curve_df)

  assign(x = "risk_curve", value = risk_curve, envir = .GlobalEnv)
}
risk_curve
ggplot(risk_curve, aes(time / 365, surv, color = as.character(T_stage), group = as.character(T_stage))) + 
  scale_color_manual(values = c("0" = "#D81B60", "1" = "#1E88E5", "2" = "#FFC107", "3" = "#004D40"), 
                    labels = c("N0", "N1", "N2", "N3"), 
                    name = "N stage") +
  geom_step(linewidth = 1) +
  labs(x = "Time (years)", y = "Recurrence-Free Probability",
       title = "Predicted Recurrence-Free Curve for ER+ or HR+ Cancer") +
  theme_minimal(base_size = 15) +
  theme(title = element_text(size = 14, face = "bold"),
        legend.title = element_text(size = 14, face = "bold"))
ggsave("Figures/ER_PR.jpg", units = "in", width = 5, height = 9, dpi = 300)

a <- data %>%
  filter(!is.na(Surgery)) %>%
  count() 

b <- data %>%
  filter(Surgery == 0) %>%
  count()

a-nrow(data)

b-nrow(data)
  
```

### risk curve 3

```{r}
str(person)
risk_curve = NULL
for(i in 0:3) {
  val <- as.character(i)
  
  person <- data.frame(
  mole_type_three        = "HER2+",                    # "HER+", "Luminal-like", "Triple Negative"
  Surgery_type         = "0",                               # levels are "0","1"; so keep it as a string
  N_stage              = i,                                 # levels are "0","1","2","3"
  Antiher2             = "Neoadjuvant Only",                # Adjuvant Only Both Neoadjuvant Only None
  Endocrine            = "None",                            # Adjuvant Only, Both, Neoadjuvant Only, None
  Oophorectomy         = "No",                              # "No" / "Yes"
  Pathologic_response  = "Not Complete",                    # Not Complete, DCIS Only, LCIS Only, Complete
  Patho_N_binary       = "medium/heavy lymph involvment",   # "No/small lymph involvment", "medium/heavy lymph involvment"
  Near_response_strict = "Not complete",                    # "Not complete" "Complete" "Near-complete"
  stringsAsFactors = FALSE
)
  
for (v in names(person)) {
  person[[v]] <- factor(person[[v]],
                        levels = levels(mod$xvar[[v]]))
}
  pred <- predict(mod, newdata = person)
curve_df <- data.frame(time = pred$time.interest,
                       surv = pred$survival[1, ])
curve_df$T_stage <- i

  name <- paste0("person_", i)
  risk_curve <- rbind2(risk_curve, curve_df)

  assign(x = "risk_curve", value = risk_curve, envir = .GlobalEnv)
}
risk_curve
ggplot(risk_curve, aes(time / 365, surv, color = as.character(T_stage), group = as.character(T_stage))) + 
  scale_color_manual(values = c("0" = "#D81B60", "1" = "#1E88E5", "2" = "#FFC107", "3" = "#004D40"), 
                    labels = c("N0", "N1", "N2", "N3"), 
                    name = "N stage") +
  geom_step(linewidth = 1) +
    scale_y_continuous(limits = c(.2, 1), expand = c(0, 0)) + 
  labs(x = "Time (years)", y = "Recurrence-Free Probability",
       title = "Predicted Recurrence-Free Curve for HER2+ Cancer") +
  theme_minimal(base_size = 15) +
 theme(
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
    axis.title.x = element_text(size = 18, face = "bold"),
    axis.title.y = element_text(size = 15, face ="bold"),
    axis.text.y = element_text(size = 14, color = "black"),
    axis.text.x = element_text(size = 14, color = "black"),
    legend.text =  element_text(size = 14, color = "black"),
    legend.title =  element_text(size = 14, color = "black", face = "bold"))
ggsave("Figures/HER2+.jpg", units = "in", width = 7, height = 7, dpi = 300)
```
```{r}
bar_plot<- data.frame(
 Neoadjuvent_Models = c("Initial Survival Forest", "Final Survival forest"),
C_index = c(0.7367926, 0.7754595),
  standard_error = c(0.0043309, 0.0042373)
)#[1] 0.7367926, 0.7324617 0.7411234 
#[1] 0.7754595, [1] 0.7712222 0.7796968


ggplot(bar_plot, aes(x = reorder(Neoadjuvent_Models, -C_index), y = C_index, fill = Neoadjuvent_Models)) + 
  geom_bar(stat = "identity", width = 0.6) +
  geom_errorbar(aes(ymin = C_index - standard_error, ymax = C_index + standard_error), 
                width = .5, linewidth = .5) + ylim(0, 1) +
  coord_cartesian(ylim = c(0.5, 1)) +
  labs(
    title = "C-index Across Neoadjuvant Models",
    x = "Model",
    y = "C-index"
  ) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1),
        legend.position = "none") +
  scale_fill_brewer(palette = "Set2")
ggsave("Comparison plot.jpg", dpi = 300)
cleaned_data$Age_at_mammo
```

